\section{Problem Formulation}
\label{sec:problem}

\tothink{full problem-formulation prose; the three-axis decision rule
must be stated precisely, with the per-cell axis assignment derived
from the landscape doc.}

\subsection{Setting}

Let $\scorerclasses = \{\scorerclass_1, \ldots, \scorerclass_5\}$
be the set of \emph{scorer classes} the suite ships:
heuristic ($\scorerclass_1$), embedding similarity ($\scorerclass_2$),
LLM classifier ($\scorerclass_3$), LLM rubric ($\scorerclass_4$), and
trajectory ($\scorerclass_5$). Let
$\taskshapes = \{\taskshape_1, \ldots, \taskshape_8\}$ be the eight
canonical \emph{task shapes} catalogued in the landscape doc
\cite{blackrim_evalsuite_landscape}: single-file edit, lookup / Q\&A,
rubric writing, ADR judgment, P0 debug triage, long-form docs,
code review, threat modelling.

A \emph{cell} $\evalcell = (\scorerclass, \taskshape) \in
\scorerclasses \times \taskshapes$ is a (scorer-class, task-shape) pair.
Of the $5 \times 8 = 40$ raw cells, only the 24 catalogued in the
landscape doc Section C are meaningful intersections; the rest are
inappropriate combinations (e.g., heuristic exact-match on long-form
docs).

\subsection{Per-cell reliability axes}

The substantive deviation from prior eval-suite work is that we
\emph{do not} collapse cell reliability to a single number. Instead
each meaningful cell $\evalcell$ is assigned a primary reliability
axis from $\{\text{calibration}, \text{agreement}\}$ and is subject
to a bias-bound precondition:

\begin{itemize}[leftmargin=*]
  \item \textbf{Calibration axis} (cells C.5--C.7, C.23). Where the
        scorer emits a confidence-vs-correctness signal, the load-bearing
        constraint is $\ECE \leq \tau_{\ECE}$ (and equivalently
        a Brier-score floor). Per-cell tolerances $\caliblimit$ are
        anchored to literature priors --- ECE 0.05--0.15 for
        forced-choice classifier on Q\&A, ECE 0.05--0.10 for
        Bradley-Terry pairwise.
  \item \textbf{Agreement axis} (cells C.4, C.8--C.14, C.19--C.22).
        Where the scorer's output is intrinsically continuous /
        ordinal but no calibration baseline exists, the load-bearing
        constraint is $\kappastat \geq \tau_{\kappastat}$ (or
        Krippendorff $\alphakr$ for the multi-rater future).
        Per-cell tolerances $\agreelimit$ are anchored to
        Landis-Koch interpretation bands plus task-specific priors
        from the landscape doc.
  \item \textbf{Bias-bound precondition} ($\biasbound$, all LLM-judge
        cells). Position-bias flip-rate $\leq 0.05$ (mitigated via
        swap test); verbosity-bias correlation $|\rho(\mathrm{length},
        \score)| \leq 0.20$ on a held-out diagnostic set; self-preference
        advantage $\leq 0.05$ score-points on a cross-family probe.
        \tothink{Self-preference probe details: cite Panickssery et al.\ /
        Wataoka et al.\ for the cross-family construction.}
\end{itemize}

\tothink{enumerate the per-cell axis assignment as a table; this
crystallises the substantive content of \cref{sec:related} and feeds
\cref{sec:eval}'s gates table.}

\subsection{Three-axis no-regression}

For each cell $\evalcell$ on the calibration axis, define the
\emph{calibration-no-regression} constraint as
\begin{equation}
  \ECE_{\evalcell}(\policy_{\mathrm{new}}) - \ECE_{\evalcell}(\policy_{\mathrm{base}})
  \;\leq\; \caliblimit(\evalcell),
  \label{eq:cal-noreg}
\end{equation}
where $\policy_{\mathrm{base}}$ is the previously-shipped configuration
on the same cell and $\caliblimit(\evalcell)$ is the per-cell drift
tolerance from \cref{sec:related}. The Brier-score equivalent is
$\Brier_{\evalcell}(\policy_{\mathrm{new}}) -
\Brier_{\evalcell}(\policy_{\mathrm{base}}) \leq \caliblimit'$.

For each cell $\evalcell$ on the agreement axis, define the
\emph{agreement-no-regression} constraint as
\begin{equation}
  \kappastat_{\evalcell}(\policy_{\mathrm{new}}) -
  \kappastat_{\evalcell}(\policy_{\mathrm{base}})
  \;\geq\; -\agreelimit(\evalcell).
  \label{eq:agree-noreg}
\end{equation}

The bias-bound precondition is asymmetric: a violation \emph{disqualifies}
$\policy_{\mathrm{new}}$ regardless of \cref{eq:cal-noreg,eq:agree-noreg}
holding. \tothink{state this precisely as: $\biasbound_{\evalcell}
(\policy_{\mathrm{new}}) \leq \tau_{\biasbound}$ where $\biasbound$
is the worst-of-three (position, verbosity, self-preference) bias
diagnostic.}

\subsection{Why three axes do not collapse}

\tothink{The case for non-collapsibility, with the failure-mode
catalogue from the landscape doc Section D. (i) Heuristic scorers
are perfectly calibrated against the oracle but can have
$\kappastat < 1$ against humans (Goodhart). (ii) Rubric scorers
can have stable $\kappastat$ across runs while $\ECE$ blows up
(self-inconsistency dominates). (iii) Verbosity-biased judges can
have $\ECE$ near zero on length-controlled subsets and arbitrarily
large on the full distribution. The frame must allow all three
to fire independently --- collapsing to a single composite hides
the failure mode operators most need to see.}

\subsection{What the eval-suite must produce}

Operationally the suite emits, per (run, cell), a triple
$(\ECE, \kappastat, \biasbound)$ plus the per-axis no-regression
verdict against $\policy_{\mathrm{base}}$. The gate is the conjunction:
ship $\policy_{\mathrm{new}}$ iff every cell's relevant axes pass
\emph{and} the bias-bound precondition holds on every LLM-judge cell.
\tothink{state the gate as a Boolean expression over per-cell
predicates; this is the core operational artefact.}
