\section{Discussion}
\label{sec:discussion}

\tothink{full discussion prose; outline below covers four sub-themes,
of which the first --- the RLHF calibration finding --- is the
substantive methodological contribution and deserves the most space.}

\subsection{The RLHF-degrades-judge-calibration finding}

\tothink{the headline empirical finding from \cref{sec:eval}, traced
to the literature ancestor. Kadavath et al.\ \cite{kadavath2022language}
documented this on multiple-choice / Q\&A tasks across the Anthropic
model line; we reproduce on judge tasks specifically. Mechanism (per
the InstructGPT line): RLHF rewards models for confident, declarative
answers, which decouples reported probability from correctness.
Implication for any team adopting LLM-as-judge: the dominant deployed
configuration --- forced-tool-use on an instruction-tuned model with
log-probs read directly --- is the configuration the literature
already documents as miscalibrated. Temperature scaling is the
cheap fix; not shipping it is the unforced error.}

\subsection{Why three axes do not collapse to one}

\tothink{the case for three axes vs a single composite. The
landscape doc Section D's per-cell axis assignment shows that
collapsing forfeits the operationally-distinguishable failure modes:
heuristic scorers fail on dataset bias (Goodhart) which is
upstream-of-scorer; rubric scorers fail on self-inconsistency
(\emph{agreement} stable, \emph{calibration} broken); verbosity-biased
judges fail on bias-bound while calibration on length-controlled
subsets looks fine. Each requires a different fix. A composite
average obscures the fix.}

\subsection{The Bruch-style open question for aggregators}

\tothink{the eval-suite analogue to the retriever paper's RRF-vs-CC
open question is whether mean-of-means, geometric-mean, or
inverse-variance weighting should be the default aggregator. The
algorithm survey scores them comparably (28 / 26 / 26 in Section B);
the decision turns on per-suite scorer-variance heterogeneity.
\cref{sec:eval} reports per-scorer $\sigma_i$ on stable suites
and the MSE gap of weighted vs flat. Geometric-mean is the right
default if the gating semantics are "must pass most scorers";
inverse-variance is the right default if scorer reliability genuinely
varies; flat mean is the right default if the suite is
hand-curated to homogeneous scorer reliability. Most operators run
heterogeneous suites and don't know it.}

\subsection{When this won't work}

\tothink{three regimes break the architecture as currently described:
(i) \emph{very small calibration sets} ($N < 50$) where ECE is
itself badly estimated (binning noise dominates) --- temperature
scaling becomes unstable; the right move at small $N$ is to use
Brier directly (no binning) and refresh more aggressively.
(ii) \emph{judge model drift} --- if the judge model is updated
silently (Anthropic ships a new haiku build), the previously-fit
$T$ is wrong; OQ-E7 governs the refresh cadence but the underlying
problem is that judges should be pinned by hash, not by model name.
(iii) \emph{rubric drift} --- if the rubric is hand-edited between
runs without re-fitting $T$, the calibration set's coverage is
invalidated; defensive move is to derive a rubric-content hash
and invalidate the calibration table on hash change.}

\subsection{Lessons from deployment}

\tothink{three deployment lessons cross-applicable beyond this paper:
(i) probability-weighted scoring is a free upgrade that no published
eval-tool ships by default --- the API exposes the log-probs and
none of OpenAI Evals, Promptfoo, Inspect, DeepEval, or Confident-AI's
G-Eval implementation reads them. Surface this as a one-paragraph
"if you ship one thing from this paper" recommendation. (ii) The
three-axis frame's bias-bound axis is the load-bearing one
operationally --- a cell that fails calibration can be fixed by
scaling; a cell that fails agreement can be fixed by rubric
sharpening; a cell that fails bias-bound is unfixable downstream.
(iii) Per-judge calibration sets are a versioned artefact like the
suites themselves. Treat them as first-class. Cite the analogous
artefact-versioning lesson from the model-advisor paper
\cite{blackrim_moa1c}.}
