\section{Notation Reference}
\label{app:notation}

\begin{center}
\begin{tabular}{ll}
\toprule
Symbol & Meaning \\
\midrule
$\fileset$, $\file$            & Set of source files; a source file \\
$\langset$, $\lang$            & Set of supported languages; a language \\
$\loc(\file)$                  & Lines of code in $\file$ \\
$\tokensize(\file)$                 & Token count of $\file$ under the LLM tokenizer \\
$\outline(\file)$              & Canonical outline of $\file$ \\
$\ratio(\file)$                & Compression ratio $|\outline(\file)|/\tokensize(\file)$ \\
$\budget$                      & Target token budget per outline (default $300$) \\
$\threshold$                   & LoC threshold for outline-required (default $200$) \\
$\intent$, $\intentset$        & A transform intent; the registered intent set \\
$\plan$, $\diff$               & A symbolic-surgery plan; its candidate diff \\
$\gate$, $\verdict$            & The compile gate; its verdict $\in \{\text{accept},\text{reject}\}$ \\
$\agent$, $\task$              & An agent; a task \\
$\success(\task; \agent)$      & Task-success indicator $\in \{0, 1\}$ \\
$\cost_\text{read}$            & Input-token cost of agent reading activity \\
\bottomrule
\end{tabular}
\end{center}

\section{Per-Language Significance Kinds}
\label{app:significance}

The hybrid significance classifier in \cref{sec:algorithm:significance}
uses a per-language fixed list (below) plus a universal fallback for
unrecognised constructs that are named or have line-span $\geq 10$.

\begin{center}
\begin{tabular}{p{2.2cm}p{10cm}}
\toprule
Language & Fixed-list significant kinds \\
\midrule
\textbf{Go} & named anonymous struct or interface; \texttt{go func()} goroutine
              spawn; \texttt{defer} blocks $\geq 3$ lines; inner
              \texttt{select} or \texttt{switch} with $\geq 3$ cases; type
              assertions on interface values. \\
\textbf{Python} & nested \texttt{def} and \texttt{class}; \texttt{@dataclass}
              and \texttt{@dataclasses.dataclass}; \texttt{with} blocks
              managing files / locks / contexts; \texttt{async def} and
              \texttt{await} constructs; decorators with non-trivial
              bodies. \\
\textbf{TypeScript / JavaScript} & nested function or arrow declarations
              bound to identifiers; React component definitions (JSX root);
              JSX trees $\geq 10$ lines; \texttt{useEffect} /
              \texttt{useMemo} / \texttt{useCallback} blocks;
              \texttt{try / catch} with $\geq 3$-line catch. \\
\textbf{Universal} & any named or labelled inner construct; any unnamed
              construct with line-span $\geq 10$. \\
\bottomrule
\end{tabular}
\end{center}

\section{Outline Markdown Schema (Specimen)}
\label{app:schema}

The canonical Markdown schema, demonstrated on a hypothetical file
\texttt{state.go} (847 LoC, package \texttt{state}):

\begin{lstlisting}[basicstyle=\ttfamily\small,frame=single]
# state.go (847 LoC, 14 decls, package `state`)

> Package state implements user session storage backed by Redis.
> Concurrency: all exported functions are safe for parallel use.
> (truncated at 3 lines)

## Imports
fmt, os, sync, internal/foo, github.com/x/y

## Types
- `User` struct (L23-L45) - fields: ID, Name, Email
- `Repository` interface (L48-L52) - Save(u User) error

## Functions
- `NewUser(name, email string) *User` (L60-L78)
  - go: cleanup goroutine (L66-L70)
  - defer: log.Close (L75)
- `(r *userRepo) Save(u User) error` (L82-L120)
  - switch: u.Status / 4 cases (L88-L105)
- `validate(u User) error` *(private)* (L125-L140)
\end{lstlisting}

The schema is line-span-anchored: every declaration carries
$L_{\text{start}}\text{-}L_{\text{end}}$, allowing an agent to issue a
targeted \texttt{Read --offset N --limit M} to fetch the body of a
specific decl without re-reading the file.

\section{Sanitisation Pattern Set}
\label{app:sanitisation}

The literal-string patterns flagged by \cref{sec:algorithm:sanitization}
as prompt-injection candidates:

\begin{center}
\small
\begin{tabular}{ll}
\toprule
Pattern & Rationale \\
\midrule
\texttt{IGNORE PREVIOUS}            & Classic prompt-injection prefix \\
\texttt{IGNORE ALL PREVIOUS}        & Variant \\
\texttt{SYSTEM:}                    & Role-impersonation marker \\
\texttt{ASSISTANT:}                 & Role-impersonation marker \\
\texttt{USER:}                      & Role-impersonation marker \\
\texttt{<|im\_start|>}              & ChatML control token \\
\texttt{<|im\_end|>}                & ChatML control token \\
\texttt{<s>}, \texttt{</s>}         & LLaMA-family control tokens \\
\texttt{[INST]}, \texttt{[/INST]}   & Instruct-tuned model markers \\
\texttt{Human:}, \texttt{Assistant:} & Anthropic-style role markers \\
\bottomrule
\end{tabular}
\end{center}

The set is conservative-biased: legitimate prose containing any of
these literals is flagged with the \texttt{[sanitised]} prefix; the
content is not dropped, only marked. Operators can extend the set in
\texttt{ship.toml [outline.sanitisation] additional\_patterns}.

\section{Pre-Registered Open Questions}
\label{app:open-questions}

The empirical questions whose answers gate promotion or demotion of
the various subsystems. Pre-registered in this paper draft and in
\texttt{docs/research/blackrim-ast-open-questions.md} in the
companion repository.

\begin{itemize}[leftmargin=*]
  \item \textbf{OQ-AST-1 — Token-budget conformance.} On the
        broader polyglot corpus, what fraction of outlines satisfy
        $|\outline(\file)| \leq \budget = 300$? Breakdown by
        language; report $p_{50}$, $p_{95}$, $p_{99}$.
  \item \textbf{OQ-AST-2 — Emission latency.} Wall-clock latency
        of \texttt{gt outline} on files 100~LoC to 5000~LoC; what
        is the user-perceived latency added by the
        prepend-to-Read hook?
  \item \textbf{OQ-AST-3 — Discipline adoption trajectory.}
        Hit-rate of outline-discipline over the 14-day rollout;
        does adoption cross 80\% triggering promote-to-block?
  \item \textbf{OQ-AST-4 — Compile-gate false-negative rate.}
        On a curated 200-diff corpus, what is the false-negative
        rate of $\gate$ (rejected diffs that ground-truth review
        would have accepted)? Is it within the 5\% conservatism
        budget?
  \item \textbf{OQ-AST-5 — End-to-end token reduction.} On paired
        Explore-tier traces (with and without discipline), what
        is the per-task input-token reduction, the wall-clock
        change, and the task-success delta?
  \item \textbf{OQ-AST-6 — Pattern-DSL authoring fidelity.} When
        the YAML pattern-DSL is implemented, what is the fidelity
        loss vs. native parser walking, measured by per-intent
        match coverage on a sample-of-intents harness?
  \item \textbf{OQ-AST-7 — Cross-file outline composition.} For
        multi-file refactors, does a cross-file outline composition
        (call graph + type dependency summary) improve agent
        success rate over per-file outlines plus
        \texttt{codeindex\_explore} retrieval?
\end{itemize}

\section{Reproducibility Checklist}
\label{app:repro}

Following the recommendations of \cite{pineau2021improving} for
empirical reproducibility:

\begin{itemize}[leftmargin=*]
  \item \textbf{Code.} The reference implementation is the Blackrim
        repository at the tagged commit cited in
        \cite{blackrim_paper_repo}. Relevant directories:
        \texttt{cmd/gt/outline/}, \texttt{cmd/gt/compress\_structure*},
        \texttt{internal/codeindex/refactor/},
        \texttt{internal/codeindex/symbol/}, \texttt{hooks-staging/}.
  \item \textbf{Data.} The benchmark fixture is in
        \texttt{cmd/gt/testdata/bench/}; the broader corpus is pinned
        in \texttt{scripts/corpus.txt}. Telemetry inputs to OQ-AST-3
        and OQ-AST-5 are released anonymised under the same tag.
  \item \textbf{Hyperparameters.} Default $\budget = 300$ tokens;
        $\threshold = 200$ LoC; sanitisation pattern set per
        \cref{app:sanitisation}. Defaults match in-tree values; both
        are per-project overridable in \texttt{ship.toml}.
  \item \textbf{Random seeds.} Outline emission and the compile gate
        are deterministic. Where any stochastic harness is used
        (e.g., paired-trace sampling for OQ-AST-5), seeds are
        published in the artefact \texttt{README}.
  \item \textbf{Compute.} The full §6.2 reproduction runs on a
        single laptop in $< 1$~minute; no GPU required.
  \item \textbf{Hardware.} Reference deployment runs on
        \texttt{darwin} and \texttt{linux}; no special hardware. The
        TypeScript backend's first-run \texttt{npm install} requires
        network access.
\end{itemize}
