\begin{abstract}
\noindent
\tothink{abstract}

The abstract will compress the paper into a single dense paragraph; outline of
the five mandatory bullet points it must hit:

\begin{itemize}[leftmargin=*]
  \item \textbf{Problem.} Multi-agent memory pools are heterogeneous (short
        bd-memory entries, mixed-shape queries, recency-sensitive); the
        retriever sits on the hot path of every agent spawn and a wrong
        recall poisons every downstream prompt. No single scorer or fusion
        configuration dominates across the six query classes Blackrim
        observes (technical-lookup, failure-recall, agent-scoped, continuity,
        concept-bridge, exact-id-or-slug).
  \item \textbf{Framing.} Selection of fusion configuration is a
        \emph{conservative contextual bandit} problem — arms are
        fusion-config combinations, context is the query class (and ideally
        a query-class classifier), reward is per-query NDCG@10, and the
        safety constraint is per-class no-regression below the BM25 (or
        BM25+decay) baseline. This is structurally distinct from the
        model-advisor paper's tier-cost-ratio framing because the cost axis
        is local latency, not LLM-token cost.
  \item \textbf{Algorithm.} The \emph{proposed} system stacks BM25 +
        SPLADE + dense + depgraph + RRF + decay + MMR + optional
        cross-encoder reranker, gated by a SPLADE corpus-size threshold
        and a Critical-tier reranker hook. The contextual-bandit policy
        on top selects per-class fusion weights subject to a
        conservative-bandit \cite{wu2016conservative,kazerouni2017conservative}
        constraint. The \emph{current prototype} ships two of these
        scorers (keyword and depgraph); BM25, SPLADE, and dense are
        implemented in-tree but not yet deployed.
  \item \textbf{Evaluation.} Empirical evaluation uses paper-stream
        telemetry (\texttt{internal/bdmemory/paper\_stream.go}) flowing to
        \texttt{data/raw/queries.jsonl}, with held-out per-class NDCG@10
        and recall@k against a labeled relevance set. Measured baselines
        use the current 2-scorer prototype (keyword, depgraph); baselines
        requiring BM25, SPLADE, or dense are deferred pending deployment
        of the proposed full stack.
  \item \textbf{Finding.} \tothink{headline finding once data accumulates --
        per-class no-regression preserved at $\budget = 0$ over the
        prototype 2-scorer arm set; total NDCG@10 improvement of
        $\tothink{X\%}$ over keyword baseline; SPLADE corpus-size gate
        expected to trigger at $\tothink{N}$ documents once deployed.}
\end{itemize}

\textbf{Keywords:} hybrid retrieval, contextual bandits, conservative
exploration, agent memory, BM25, SPLADE, RRF, recency decay, NDCG,
depgraph retrieval.
\end{abstract}
