\section{Introduction}
\label{sec:intro}

\tothink{full introduction prose; sequenced after model-advisor paper.}

This section will develop:

\begin{itemize}[leftmargin=*]
  \item \textbf{The retriever as the hot path.} Every agent spawn starts
        with a \texttt{gt context recall} or \texttt{gt context budget}
        call into \bdmemory; the retrieved memories enter the prompt
        prefix that determines cache-hit behaviour. A wrong recall poisons
        every downstream prompt; a slow recall blocks dispatch. The
        algorithmic choice on this path has compounding effects.
  \item \textbf{Heterogeneous memory pool.} The bd-memory pool is
        characterised by short documents (typically 30--400 tokens),
        agent-prefix tagging (\texttt{[Yatima]}, \texttt{[Blanca]}),
        \texttt{[failure]}-prefix structural priors, and recency-sensitive
        engineering knowledge that ages quickly. No single retriever
        family dominates across this surface --- \tothink{cite landscape
        doc Section A six query classes (Q1--Q6)}.
  \item \textbf{Why no single configuration wins.} A short summary of
        the per-(query-class $\times$ scorer-family) matrix in the
        landscape doc: BM25 dominates Q6 (exact-ID), hybrid-RRF wins on
        Q5 (concept-bridge), recency-decay is load-bearing on Q4
        (continuity). The "average NDCG" framing hides the per-class
        regression risk that operators care about.
  \item \textbf{Framing.} Why \emph{conservative contextual bandits}
        \cite{wu2016conservative,kazerouni2017conservative} is the right
        formalism here, and why this is methodologically distinct from
        MOA-1b's tier-cost-ratio framing (cost axis is local latency,
        not LLM tokens; baseline is per-class BM25, not a single
        agent-default tier).
  \item \textbf{Contributions.} (i) Per-(query-class $\times$ scorer-family
        $\times$ fusion-config) priors grounded in IR literature
        (\cref{sec:related,sec:problem}); (ii) a fusion architecture
        \cref{sec:system} stacking BM25/SPLADE/dense/RRF/decay/MMR/rerank
        with explicit corpus-size and tolerance-class gates; (iii) a
        conservative-contextual-bandit policy \cref{sec:algorithm}
        with per-class baselines as the safety constraint; (iv) an
        empirical evaluation pipeline \cref{sec:eval} backed by
        \texttt{internal/bdmemory/paper\_stream.go} telemetry.
\end{itemize}
