\section{Problem Formulation}
\label{sec:problem}

\tothink{full problem-formulation prose; the decision rule and the
conservative-bandit constraint must be stated precisely.}

\subsection{Setting}

Let $\queryclasses = \{\queryclass_1, \ldots, \queryclass_{|\queryclasses|}\}$
be a finite set of \emph{query classes}. In \blackrim's deployment,
$|\queryclasses| = 6$ corresponding to the classes derived in the landscape
doc \cite{blackrim_ijl0_landscape} from inspection of the recall call
sites: technical-lookup ($\queryclass_1$), failure-recall ($\queryclass_2$),
agent-scoped ($\queryclass_3$), continuity ($\queryclass_4$),
concept-bridge ($\queryclass_5$), exact-id-or-slug ($\queryclass_6$).

Let $\armset = \{\arm_1, \ldots, \arm_{|\armset|}\}$ be a finite set of
\emph{fusion-config arms}. Each arm $\arm$ is a tuple specifying a
first-stage scorer set $\scorerset(\arm) \subseteq \scorers$, a fusion
operator, optional decay/MMR/rerank wrappers, and the per-component
weights.

The scorer vocabulary $\scorers$ is deployment-dependent.  In the
current production prototype (\S\ref{sec:system-prototype}):
\[
  \scorers_{\text{prototype}} = \{\texttt{keyword},\; \texttt{depgraph}\}.
\]
In the proposed full stack (\S\ref{sec:system-proposed}):
\[
  \scorers_{\text{proposed}} = \{\texttt{BM25},\; \texttt{SPLADE},\;
                                  \texttt{dense},\; \texttt{depgraph}\}.
\]

The bandit problem formulation, arm enumeration, and conservative
constraint below hold for \emph{any} finite $\scorers$; the two
concrete vocabularies are instances.  Quantitative results in
\cref{sec:eval} use $\scorers_{\text{prototype}}$ (the set available
from paper-stream telemetry at the time of writing); results against
$\scorers_{\text{proposed}}$ are deferred pending the deployment of
SPLADE and dense retrieval.

\tothink{enumerate the canonical arms for the prototype (keyword-only;
keyword+decay; keyword+depgraph-RRF; keyword+depgraph-RRF+decay) and for
the proposed stack (BM25-only; BM25+decay; hybrid-RRF (BM25+SPLADE);
hybrid-RRF (BM25+dense); hybrid-RRF (BM25+SPLADE+dense); hybrid-CC
(BM25+dense, $\alpha$ tuned); hybrid-RRF + decay; hybrid-RRF + MMR;
hybrid-RRF + rerank).}

\subsection{Reward}

For each query $x$ classified into $\queryclass(x) \in \queryclasses$ and
each arm $\arm \in \armset$, the policy emits a top-$k$ ranked list of
documents. The realised reward $\reward(x, \arm) \in [0, 1]$ is per-query
NDCG@10 against a held-out relevance judgement. Aggregate reward over a
trace $\{x_1, \ldots, x_T\}$ is the average per-query NDCG@10:
\(
  R_T(\policy) = \tfrac{1}{T} \sum_{t=1}^T \reward(x_t, \policy(x_t)).
\)

We use NDCG@10 as the primary metric; MRR@10 and recall@k are reported as
secondary. \tothink{justify NDCG@10 over alternatives; cite Smucker et al.
2007 \cite{smucker2007comparison} on significance testing for IR metrics.}

\subsection{Per-class baselines and the safety constraint}

The methodological framing follows the conservative-contextual-bandit
literature \cite{wu2016conservative,kazerouni2017conservative}. For each
query class $\queryclass \in \queryclasses$, let $\baseline(\queryclass)$
denote the per-class \emph{baseline arm} the policy must not regress
below:

\begin{itemize}[leftmargin=*]
  \item $\baseline(\queryclass_1, \queryclass_3, \queryclass_5, \queryclass_6) = $
        BM25-only (proposed stack); keyword-only (prototype).
  \item $\baseline(\queryclass_4) = $ BM25+decay (proposed stack);
        keyword+decay (prototype).
  \item $\baseline(\queryclass_2) = $ keyword + failure-tag boost
        (both stacks; this scorer ships in the prototype).
\end{itemize}

The prototype baselines use \texttt{keyword} where the proposed stack
uses BM25 because BM25 is not yet deployed (see \S\ref{sec:system}).
Empirical baseline measurements in \cref{sec:eval} use the prototype
vocabulary.  The choice of per-class baseline is the substantive
deviation from the model-advisor paper's tier-cost-ratio framing
\cite{blackrim_moa1c}. The conservative-bandit constraint at confidence
level $1 - \alpha$ (we use $\alpha = 0.05$) is
\begin{equation}
  \sum_{t \le T : \queryclass(x_t) = \queryclass}
    \big[\reward(x_t, \policy(x_t)) - \reward(x_t, \baseline(\queryclass))\big]
  \;\geq\; -\budget,
  \label{eq:conservative-budget}
\end{equation}
with high probability uniformly over $T$, where $\budget \geq 0$ is the
\emph{conservative budget}. \tothink{tighten the high-probability
statement to match Wu et al.\ Theorem 2 / Kazerouni et al.\ Theorem 1.}

\Cref{eq:conservative-budget} encodes "no per-class regression below
baseline." Critical-tier queries (\tothink{define which classes count as
Critical in deployment --- likely $\queryclass_2$ failure-recall and
$\queryclass_6$ exact-ID}) get $\budget = 0$ (hard no-regression);
non-Critical classes get small positive $\budget$ for exploration headroom.

\subsection{The bandit problem}

The advisor's job is to learn a policy $\policy : \queryclasses \to \armset$
that maximises expected NDCG@10 subject to \cref{eq:conservative-budget}.
This is a contextual-bandit problem with discrete arms, side-information
(the query class), and a per-context-conditional safety budget.
\tothink{state the formal regret-vs-baseline bound; cite Wu et al.\ /
Kazerouni et al.\ for the high-probability conservative-bandit guarantee.}

\subsection{What the advisor must produce}

Operationally, the advisor exposes a function
\(
  \mathrm{advise} : (x, \queryclass(x)) \mapsto (\arm^\dagger, \mathrm{rationale})
\)
returning the recommended arm and a structured rationale (per-arm
posterior mean reward, conformal lower bound, per-class baseline
comparison). The rationale is non-optional for the same audit reasons as
in \cite{blackrim_moa1c}: operators must be able to override and explain.
