\section{Algorithm: Conservative Contextual Bandit over Fusion Configs}
\label{sec:algorithm}

\tothink{full algorithm prose; the conservative-contextual-bandit policy
specification is the load-bearing piece.}

\subsection{Layered architecture}

The proposed policy stacks four layers, in the same shape as MOA-1c's
CC-TS but adapted to the retrieval surface:

\begin{description}
  \item[Layer 1: Per-(class, arm) reward posteriors.] Beta posteriors
        on per-query NDCG@10, with offline-prior pseudocounts seeded
        from the per-(query-class $\times$ scorer-family $\times$
        fusion-config) matrix in the landscape doc
        \cite{blackrim_ijl0_landscape}.
  \item[Layer 2: Conservative gate.] The conservative-bandit constraint
        \cref{eq:conservative-budget} filters arms whose per-class
        cumulative regret against $\baseline(\queryclass)$ would breach
        the budget at confidence $1 - \alpha$.
  \item[Layer 3: Posterior-sample selection.] Among admitted arms,
        Thompson Sampling \cite{russo2018tutorial} on the posterior over
        per-class arm rewards.
  \item[Layer 4: Trigger-based eval probes.] When a per-(class, arm)
        posterior is too wide for Layer 2 to commit, schedule a labeled
        eval probe to tighten the posterior --- the same eval-trigger
        loop used in CC-TS.
\end{description}

\subsection{Layer 1: Beta posteriors over NDCG@10}

\tothink{justify Beta as the right family for $\reward \in [0,1]$;
discounted Thompson sampling for non-stationarity \cite{rajcoomar2017discounted}
parking-lot.}

\paragraph{Cold-start prior.}
\tothink{seed prior from landscape doc Section C confidence percentages
exactly as in MOA-1c eq:prior-from-confidence: $c \mapsto (\max(1,
\round(c/5)), \max(1, \round((100-c)/5)))$. The matrix in the landscape
doc gives a confidence per (class, scorer) cell; we lift those to
per-(class, arm) confidences by combining the per-component
confidences for the arm.}

\subsection{Layer 2: Conservative gate}

\tothink{state the per-class cumulative-regret budget tracker; the gate
admits arm $\arm$ for class $\queryclass$ at time $t$ only if the
expected regret-against-baseline up to $t$, plus the worst-case regret
of selecting $\arm$ now, stays within $-\budget(\queryclass)$ at
confidence $1 - \alpha$. Closed-form via per-(class, arm) posterior
means and conformal bounds.}

\subsection{Layer 3: Posterior-sample selection}

\tothink{Among gate-admitted arms, draw a sample
$\widehat{r}(\arm, \queryclass) \sim \posterior_{\queryclass, \arm}$ for
each, pick $\argmax$. Latency-cost tiebreak when posterior samples are
within $\epsilon$ of each other --- prefer the cheaper arm.}

\subsection{Layer 4: Trigger-based eval probes}

\tothink{When the posterior $\posterior_{\queryclass, \arm}$ has
credibility-interval width $> \theta_\mathrm{eval}$, schedule a labeled
eval probe (paired query against held-out relevance judgement). The
labeled-probe budget per session is operator-configurable; default
$5$ probes/day.}

\subsection{The CCBandit-Retriever algorithm}

\Cref{alg:ccbandit} states the policy in full.

\begin{algorithm}[h]
\KwIn{Query $x$ with class $\queryclass(x) = \queryclass$,
      per-(class, arm) posteriors $\posterior$, baseline arms
      $\baseline(\cdot)$, budget $\budget(\cdot)$,
      cumulative regret tracker $G(\cdot, \cdot)$, confidence
      $\alpha = 0.05$.}
\KwOut{Recommended arm $\arm^\dagger$, rationale $r$.}

\BlankLine
$\mathrm{cands} \leftarrow \{\baseline(\queryclass)\}$
  \tcp*{baseline always admitted}
\For{$\arm \in \armset \setminus \{\baseline(\queryclass)\}$}{
  $g_\mathrm{worst} \leftarrow \mathrm{WorstCaseRegret}(\arm, \queryclass,
       \posterior_{\queryclass, \arm}, \alpha)$\;
  \If{$G(\queryclass, \arm) + g_\mathrm{worst} \geq -\budget(\queryclass)$}{
    $\mathrm{cands} \leftarrow \mathrm{cands} \cup \{\arm\}$
      \tcp*{Layer 2 admits}
  }
}

\For{$\arm \in \mathrm{cands}$}{
  $\widehat{r}(\arm) \leftarrow $ sample from $\posterior_{\queryclass, \arm}$
  \tcp*{Layer 3 Thompson sampling}
}
$\arm^\dagger \leftarrow \argmax_{\arm \in \mathrm{cands}} \widehat{r}(\arm)$\;

\If{$\mathrm{CIWidth}(\posterior_{\queryclass, \arm^\dagger}, \alpha)
     > \theta_\mathrm{eval}$}{
  $\mathrm{ScheduleEvalProbe}(\queryclass, \arm^\dagger)$
    \tcp*{Layer 4 trigger}
}

\Return{$\arm^\dagger,
  r = \{\mathrm{cands}, \widehat{r}, G(\queryclass, \cdot)\}$}\;

\caption{CCBandit-Retriever recommend.}
\label{alg:ccbandit}
\end{algorithm}

\subsection{Theoretical guarantees}

\tothink{state the conservative-bandit regret bound. Two results:
(i) regret vs.\ optimal-arm-per-class scales as
$\mathcal{O}(\sqrt{|\armset| T \log T})$ via Wu et al.\ Theorem 2;
(ii) per-class no-regression-vs-baseline holds with probability
$\geq 1 - \alpha$ uniformly in $T$ via the conservative budget. Both
are direct corollaries of \cite{wu2016conservative,kazerouni2017conservative}
applied to the discrete-arm contextual bandit with
side-information ($\queryclass$).}

\subsection{Deployment status and prototype scope}

The CCBandit-Retriever algorithm is scorer-agnostic: \cref{alg:ccbandit}
operates over whatever arm set $\armset$ is enumerated for the active scorer
vocabulary $\scorers$ (\S\ref{sec:problem}).  The current production prototype
ships $\scorers_{\text{prototype}} = \{\texttt{keyword}, \texttt{depgraph}\}$,
giving a 2-arm base from which fusion-config arms are drawn.  The algorithm has
been deployed against this prototype in the Blackrim paper-stream telemetry
pipeline: per-recall events flowing to
\texttt{data/raw/queries.jsonl} carry both scorer contributions, and the
conservative-bandit policy selects among keyword-only, depgraph-only, and
keyword+depgraph-RRF arms per query class.

Scaling to the proposed 5-scorer vocabulary
$\scorers_{\text{proposed}} = \{\texttt{BM25}, \texttt{SPLADE}, \texttt{dense},
\texttt{depgraph}\}$ requires shipping SPLADE and dense retrieval into
\texttt{internal/bdmemory/}; no changes to the algorithm are needed.  All
quantitative results in \cref{sec:eval} are measured against the
prototype vocabulary; results against the proposed stack are marked
\tothink{} pending that deployment.

\subsection{Implementation notes}

\tothink{the policy lives at \texttt{cmd/gt/contextrecall/cb\_policy.go}
(planned). The function $\mathrm{Recommend}(x, \queryclass) \to (\arm^\dagger,
r)$ is pure; per-class regret-tracker state is a separate persistent struct.
Determinism for tests: the Beta sampler is seedable.}
