\begin{abstract}
\noindent
Multi-agent workflows built on large language models (LLMs) increasingly route
work across heterogeneous model tiers --- a small fast model for lookups, a
mid-sized model for routine implementation, a frontier model for judgment
calls. Selecting the cheapest model that preserves task quality is the central
cost-control problem in production agent systems, but it sits in a regime that
existing online-learning theory addresses awkwardly: rewards are
quality-vs-cost trade-offs that are asymmetric (a single poor decision in a
critical-path task can erase weeks of cost savings), per-cell sample sizes are
small (most agent--shape combinations see fewer than ten dispatches per day),
and a meaningful Bayesian prior is available from offline evaluation but is
itself imperfect.

We formalise the problem as a constrained Bayesian decision problem: choose
the cost-minimal tier whose posterior probability of preserving quality
exceeds a per-task tolerance $\qtol$. We instantiate this with
\emph{Conservative Constrained Thompson Sampling} (CC-TS): hierarchical Beta
priors seeded from an empirical landscape of $17 \times 5 \times 3$
agent--shape--tier cells; per-cell Beta-Bernoulli posteriors updated from
production telemetry; conformal-prediction wrappers that bound quality drop
at a chosen confidence level; uncertainty-triggered eval to resolve
underdetermined cells. The conservative property --- never downgrade unless
the posterior credibly exceeds tolerance --- is preserved through a one-sided
hypothesis-testing rule with explicit asymmetric loss.

We deploy CC-TS in \blackrim, an open-source multi-agent framework
for the Claude Code runtime, across a target grid of $17$ agents and
$\approx{170}$ task cells. Over $335$ production dispatches (across $3$
active cells) and a held-out eval suite of $96$ fixtures, CC-TS yields
$79.8\%$ end-to-end cost reduction versus an opus-default policy
($\$2.03$ vs.\ $\$10.05$ at illustrative prices, see \cref{sec:eval});
on this trace, where $333/335$ dispatches land in a single high-confidence
cell, CC-TS and a static-frontmatter dispatcher coincide on cost, with
the differentiator (active downgrade of thin-evidence cells) materialising
only as the cell grid populates. Empirical quality regression on the
eval suite is $+3.12\,\mathrm{pp}$ ($95\%$ bootstrap CI
$[-2.08,\,+10.42]\,\mathrm{pp}$), within bootstrap error of opus-default. We discuss limitations --- in particular, the off-policy
evaluation gap when an agent's dispatch policy and reward distribution are
both non-stationary --- and propose a hierarchical-bandit extension for
multi-tenant federation.

\textbf{Keywords:} multi-armed bandits, contextual bandits, Thompson sampling,
conformal prediction, model routing, agentic AI, LLM cost optimization,
asymmetric loss.
\end{abstract}
