\section{Future Work}
\label{sec:future}

CC-TS as deployed today is a Tier-1 implementation: per-cell tabular
posteriors, closed-form conformal calibration, asymmetric-loss decision
rule. The open-questions list (\cref{sec:eval} and \cite{blackrim_moa1c})
points to a credible Tier-2 stack, plus three Tier-3 directions.

\paragraph{Tier 2: post-deployment refinement.}
\begin{itemize}[leftmargin=*]
  \item \textbf{Hierarchical partial pooling at full inference.}
        \Cref{eq:pooling} approximates Beta-Beta pooling closed-form. A
        full Stan or PyMC-based hierarchical model
        \cite{carpenter2017stan,gelman2013bayesian} would use both
        within-agent and cross-agent structure. The marginal benefit
        depends on the answer to OQ-A8 (cross-agent transfer) in
        \cite{blackrim_moa1c}.
  \item \textbf{Per-tier eval-trigger thresholds.}
        $\theta_{\mathrm{eval}}$ is uniform across tolerance classes today.
        Critical cells with thin evidence should arguably trigger eval at
        a tighter posterior width than Lenient cells; the empirical
        sweet-spot per class is OQ-A9.
  \item \textbf{Doubly-robust shadow policies.}
        Logging on-policy decisions with their propensities enables
        cleaner off-policy evaluation of speculative policy changes
        \cite{thomas2016data,jiang2016doubly}. We log this today but do not
        yet act on it; promoting it to a continuous shadow-evaluation loop
        is a clean extension.
\end{itemize}

\paragraph{Tier 3: parking lot.}
\begin{itemize}[leftmargin=*]
  \item \textbf{Contextual upgrade.} If task-shape features (token count,
        file count, time-of-day, prior-failure flag) move the optimal tier
        within a cell, the per-cell tabular posterior is misspecified. A
        promotion to LinUCB \cite{li2010contextual} or contextual
        Thompson Sampling \cite{agrawal2013thompson} is the textbook fix.
        OQ-A2 in \cite{blackrim_moa1c} resolves the empirical question.
  \item \textbf{Non-stationarity handling.} If model providers change tier
        behaviour over weeks --- model deprecations, prompt-routing
        changes --- the IID Beta-Bernoulli assumption fails. Discounted
        Thompson Sampling \cite{rajcoomar2017discounted} or EXP3-IX
        \cite{neu2015explore} are the standard responses; OQ-A1 resolves
        which.
  \item \textbf{Federated multi-tenant CC-TS.} The most ambitious
        extension. Consumer repos onboarded via \texttt{blackrim-project on}
        could pool posteriors across tenants without sharing raw
        telemetry, using federated Bayesian inference
        \cite{shi2024federated} or differential-privacy mechanisms. The
        bandwidth-vs-noise tradeoff is the open theoretical question; the
        privacy-and-trust story is the open product question.
\end{itemize}

\paragraph{Cross-cutting: continuous-quality signal.}
The Beta-Bernoulli framing depends on a binary success indicator. As eval
suites mature toward continuous quality scores ($\in [0,1]$ with
calibrated semantics), the natural posterior shape is Gaussian or
truncated-Gaussian. The decision rule and conservative gate generalise
naturally; only the prior elicitation in \cref{eq:prior-from-confidence}
needs revisiting. OQ-A6 governs the empirical case.
