\section{Conclusion}
\label{sec:conclusion}

We presented Conservative Constrained Thompson Sampling (CC-TS), a Bayesian
adaptive routing policy for multi-agent LLM workflows. The policy combines
hierarchical Beta-Bernoulli posteriors seeded from offline evaluation, a
conformal-LCB-gated conservative-rejection rule, an asymmetric-loss
decision step, and uncertainty-triggered eval. The conservative-property
guarantee (\cref{thm:conservative}) is the load-bearing design property:
no downgrade is committed unless the conformal lower bound credibly
exceeds the per-task quality tolerance. We deployed CC-TS in \blackrim
across $17$ agents and $\approx 170$ cells; the cost-quality tradeoff
sits on the Pareto frontier consistent with our design intent.

The paper's central claim is methodological: \emph{adaptive model-tier
selection in production agent systems is a constrained Bayesian decision
problem, not a contextual-bandit problem with an extra-large reward.} The
asymmetric-loss structure and the small-sample-per-cell regime push the
problem out of standard contextual-bandit territory and into the
conservative-bandit subliterature \cite{wu2016conservative}, calibrated by
conformal prediction \cite{angelopoulos2021gentle}, and warm-started by an
offline empirical landscape. We believe this combination is generally
applicable to any agent framework with heterogeneous model menus and
asymmetric task stakes; we hope the open-source artefact lowers the
barrier to adoption.

\paragraph{Acknowledgements.}
This work was developed alongside the open-source \blackrim codebase. We
thank the contributors and operators who provided the dispatch traces and
operator-survey data that backed \cref{sec:eval}. The companion artefacts
are released under permissive licenses; see \cite{blackrim_paper_repo} for
the code and data, and \cite{german2026blackrim} for the framework itself.
