\section{Notation Reference}
\label{app:notation}

\begin{center}
\begin{tabular}{ll}
\toprule
Symbol & Meaning \\
\midrule
$\agentset$, $\agent$ & Set of agents; an agent \\
$\shapeset$, $\shape$ & Set of task shapes; a shape \\
$\tierset$, $\tier$, $\tier^*$ & Set of model tiers; a tier; baseline tier \\
$\theta_{\agent,\shape,\tier}$ & Quality success probability at cell \\
$\posterior$, $\prior$ & Posterior, prior on $\theta$ \\
$\qtol$ & Quality-loss tolerance per cell \\
$\alpha$ & Posterior-credibility level (we use $0.05$) \\
$\widehat{q}_{\mathrm{lo}}$ & Conformal lower bound on $\theta$ \\
$\Delta\cost$ & Cost differential between two tiers \\
$\kappa$, $M[\cdot]$ & Asymmetric-loss multipliers per tolerance class \\
$\theta_{\mathrm{eval}}$ & Posterior-width threshold for eval triggering \\
$\lambda$ & Pooling weight cap (we use $0.5$) \\
\bottomrule
\end{tabular}
\end{center}

\section{Proof of \cref{thm:conservative} (full version)}
\label{app:proof}

We give the full version of the proof sketched in \cref{sec:algorithm}.

Fix a cell $(\agent, \shape)$ with baseline tier $\tier^*$ and tolerance
$\qtol$. Let $\tier \prec \tier^*$ be a candidate downgrade. By
\cref{alg:ccts}, $\tier$ is admitted to $\mathrm{cands}$ only if
\(
  \widehat{q}_{\mathrm{lo}}(\agent, \shape, \tier) \geq \mu^*(\agent, \shape) - \qtol.
\)

Conformal prediction \cite{vovk2005algorithmic,angelopoulos2021gentle}
guarantees that, under exchangeability of the calibration set with the
test point, the marginal coverage probability is at least $1 - \alpha$:
\[
  \Pr\!\big[
    \theta_{\agent,\shape,\tier}
      \in [\widehat{q}_{\mathrm{lo}}, \widehat{q}_{\mathrm{hi}}]
  \big]
  \;\geq\; 1 - \alpha.
\]
In particular,
\(
  \Pr[\theta_{\agent,\shape,\tier} \geq \widehat{q}_{\mathrm{lo}}]
    \geq 1 - \alpha.
\)

Combining with the gate condition,
\[
  \Pr\!\big[
    \theta_{\agent,\shape,\tier}
      \geq \mu^*(\agent, \shape) - \qtol
  \big]
  \;\geq\;
  \Pr\!\big[
    \theta_{\agent,\shape,\tier} \geq \widehat{q}_{\mathrm{lo}}
  \big]
  \;\geq\; 1 - \alpha,
\]
where the first inequality follows because the gate enforces
$\widehat{q}_{\mathrm{lo}} \geq \mu^* - \qtol$.

Approximating $\mu^*$ by $\theta_{\agent,\shape,\tier^*}$ (justified
asymptotically as the baseline-tier posterior concentrates), we obtain
\cref{eq:conservative}.

The Layer 3 asymmetric-loss step
\(
  \tier^\dagger = \arg\min_{\tier \in \mathrm{cands}}
    \mathbb{E}_{\theta \sim \posterior_{\tier}}
      [L(\tier, \theta; \tier^*, \mu^*)]
\)
selects from $\mathrm{cands}$; it never adds tiers, so the inequality
\cref{eq:conservative} is preserved by the final selection
$\tier^\dagger$. \qed

\section{Per-cell prior pseudocount derivation}
\label{app:prior}

The mapping
\(
  c \mapsto (\alpha_{\mathrm{prior}}, \beta_{\mathrm{prior}})
\)
in \cref{eq:prior-from-confidence} is chosen to satisfy three constraints:

\begin{enumerate}[leftmargin=*]
  \item \emph{Monotonicity in confidence.} A 95\%-confidence cell yields a
        Beta posterior whose mean exceeds a 65\%-confidence cell's mean.
  \item \emph{Bounded influence.} No prior pseudocount exceeds $20$ on
        either side, so $\sim 20$ posterior observations are enough to
        dominate even the strongest prior.
  \item \emph{Sane uniform.} A 50\%-confidence cell yields
        $\mathrm{Beta}(10, 10)$ rather than $\mathrm{Beta}(1, 1)$; this
        encodes ``we know it's roughly even'' rather than ``we know
        nothing.''
\end{enumerate}

The linear scaling $c / 5$ produces this monotone, bounded mapping with
the appropriate uniform behaviour. Alternative scalings ($c / 2$, $c / 10$)
were tested in offline simulation and produced either too-stiff or too-soft
priors; the $c / 5$ choice was the empirical sweet-spot.

\section{Selected hyperparameters}
\label{app:hyperparams}

\begin{center}
\begin{tabular}{lll}
\toprule
Hyperparameter & Value & Rationale \\
\midrule
$\alpha$ (conformal)        & $0.05$  & 95\% credibility, conventional \\
$\theta_{\mathrm{eval}}$    & $0.10$  & posterior CI half-width threshold \\
$\lambda$ (pooling cap)     & $0.5$   & max half of own-cell strength \\
$M[\text{Critical}]$        & $\infty$ & hard rejection \\
$M[\text{Strict}]$          & $20$    & ADR-cascade-equivalent \\
$M[\text{Moderate}]$        & $5$     & multi-file-ticket cascade \\
$M[\text{Lenient}]$         & $1$     & local consequence only \\
$\epsilon$ (baseline)       & $0.1$   & vanilla bandit comparator \\
DR-OPE bootstrap            & $1000$  & resamples for CIs \\
\bottomrule
\end{tabular}
\end{center}

\section{Reproducibility checklist}
\label{app:repro}

Following the recommendations of \cite{pineau2021improving} for empirical
reproducibility:

\begin{itemize}[leftmargin=*]
  \item \textbf{Code.} The reference implementation is at
        \cite{german2026blackrim} (\texttt{internal/dispatch/}). Tagged
        commit for the published version is in \cite{blackrim_paper_repo}.
  \item \textbf{Data.} The production dispatch trace is anonymised and
        released alongside the paper artefacts in \texttt{data/}; raw
        per-dispatch records are pseudonymised by agent ID hashing.
  \item \textbf{Hyperparameters.} See \cref{app:hyperparams}.
  \item \textbf{Random seeds.} All RNGs are seeded from a master seed
        published in the artefact \texttt{README}.
  \item \textbf{Compute.} The full evaluation runs on a single laptop
        in $< 1$ hour; no GPU required.
  \item \textbf{Hardware.} Production deployment runs in user terminals
        on \texttt{darwin} and \texttt{linux}; no special hardware.
\end{itemize}

\section{Baseline-policy replay}
\label{sec:baseline-replay}
\label{app:baselines}

To contextualise cc-ts performance, we replay three fixed policies
against the same 295 dispatch contexts recorded in
\texttt{data/raw/advisor-decisions.jsonl}.

\paragraph{Policies.}

\begin{description}[leftmargin=*]
  \item[static-fm.]  A static lookup table mapping agent role to tier:
        Architect $\to$ opus, Builder/Reviewer/Tester $\to$ sonnet,
        Writer/Researcher $\to$ haiku.  No per-dispatch adaptation.
        This is the implicit policy encoded in the project \texttt{CLAUDE.md}
        delegation-discipline table, applied rigidly without any
        posterior update.

  \item[$\varepsilon$-greedy ($\varepsilon = 0.10$).]  At each dispatch,
        explores (uniform random tier) with probability $\varepsilon$;
        exploits (highest Laplace-smoothed success rate within the task
        shape) with probability $1 - \varepsilon$.  Per-shape success rates
        are seeded from \texttt{data/raw/eval-triggered-observations.jsonl}
        when available.  The exploration parameter $\varepsilon = 0.10$
        matches the value in \cref{app:hyperparams}.

  \item[opus-default.]  Always dispatches to opus regardless of agent or
        task shape.  This is the conservative upper-capability baseline
        that most users implicitly apply before any advisor is in place.
\end{description}

\paragraph{Cost model.}
All cost figures use the same published Anthropic per-token prices as
\texttt{scripts/aggregate-by-shape.py}: haiku \$0.25/MTok input
(\$0.50 output), sonnet \$3.00/MTok (\$6.00), opus \$15.00/MTok
(\$30.00).  Because the per-dispatch JSONL records do not carry exact
token counts, we apply a \emph{synthetic representative token budget}:
1\,200 input tokens and 400 output tokens per dispatch.  These are
rounded illustrative figures chosen for exposition — they are \textbf{not}
measured per-call averages.  Given a fixed budget the cost numbers are
\textbf{exact} (deterministic arithmetic, no estimation).

\paragraph{Quality model (illustrative, not measured).}
cc-ts quality is normalised to $1.000$ per dispatch.  For any dispatch
where a baseline policy chose a tier \emph{below} the cc-ts recommendation
we apply a conservative $5\,\mathrm{pp}$ degradation per tier step down.
For dispatches at or above the cc-ts tier no penalty is applied.  This
worst-case model is \textbf{illustrative only}: real quality data requires
running the held-out eval fixtures (\texttt{eval-fixtures.jsonl}).
As of 2026-05-16, \texttt{data/raw/eval-fixtures.jsonl} is available
($96$ rows, $32$ fixtures $\times$ $3$ tiers; see \cref{sec:eval}).
The illustrative quality model above remains for the baseline-replay
comparison; the measured per-tier pass rates are reported in Table~\ref{tab:quality-table}.

\paragraph{Results.}

\input{figures/baseline-table}

\noindent
The replay reveals two key contrasts.  \emph{opus-default} achieves the
same quality score as cc-ts (no penalty — it always meets or exceeds the
cc-ts tier) but at nearly $5\times$ the cost (\$10.05 vs.\ \$2.03).
\emph{static-fm} happens to match cc-ts exactly on this trace because the
335 records are dominated by Builder (shape \texttt{Bu1}) dispatches that
the static table also routes to sonnet; cost savings versus opus-default
are identical.  \emph{$\varepsilon$-greedy} achieves the lowest nominal
cost (\$0.50, a $75.4\%$ reduction vs.\ cc-ts) through aggressive
exploration toward haiku, but incurs a $4.7\,\mathrm{pp}$ illustrative
quality penalty — a reminder that raw exploration without the
conservative-guarantee gate is not free.  These figures must be
interpreted with caution given the illustrative quality model; they are
presented to frame the design space, not as measured head-to-head results.
