\section{Introduction}
\label{sec:intro}

\subsection*{Cost of static-prefix multi-agent systems}

Multi-agent LLM applications that coordinate via a single orchestrator
load a canonical instruction document—typified by \claudemd in
\blackrim—into every agent spawn as a cached prefix. At operational
scale, this becomes a substantial cost driver. \blackrim executes
approximately 50 agent spawns per day across five parallel agent waves,
each loading a 708-line, 48\,487-character instruction document before
task execution begins. Anthropic's prompt cache mitigates this cost by
amortizing repeated prefixes within a 5-minute TTL window, compressing
cache hits to approximately 10\,\% of the input token cost of a cache
miss. However, TTL expiration and configuration changes invalidate
the cache and trigger full prefill charges. At observed cache hit rates
of 0.80--0.90, the effective cost per spawn remains substantial; worse,
every cache miss resets the window, creating a sawtooth cost pattern
where initial spawns pay full prefix cost and subsequent spawns within
the window pay amortized rates.

\subsection*{The compression trap}

The natural response is lossy compression. The literature offers mature
techniques: LLMLingua and LongLLMLingua aggressively compress retrieval
contexts \cite{llmlingua,longllmlingua}; RECOMP selectively augments
\cite{recomp}; AutoCompressors and ICAE train specialized models
\cite{autocompressors,icae}. All report fidelity preservation on QA and
retrieval benchmarks. However, instruction documents are structurally
different from retrieval contexts: they encode procedural rules,
dispatch logic, and governance constraints that are brittle to missing or
abbreviated context in ways QA passages are not. The critical gap is
that no prior work benchmarks lossy compression specifically on
instruction documents. The field therefore lacks measurement to assure
that compression techniques that preserve QA accuracy also preserve
instruction-following fidelity.

Meanwhile, empirical work on attention patterns offers a second reason
for caution. Liu et al.'s ``Lost in the Middle'' finding established
that models attend preferentially to early and late positions in long
contexts, with degradation in the middle. However, this U-shaped effect
is pronounced for retrieval tasks (15--25\,\% quality loss at middle
positions) but modest for instruction-tuned models (3--8\,\% only)
\cite{lostinmiddle}. Static-prefix instruction documents placed at
context start already align with the optimal attention pattern,
removing a key motivation for compression.

\subsection*{Our approach: conservative frequency-weighted disclosure}

Rather than compression, we propose frequency-weighted progressive
disclosure: partitioning the instruction document into sections and
deciding residency (in-prefix) versus externalization (on-demand Read)
based on per-section access frequency under the cost model derived in
\cref{sec:problem}. The decision rule is conservative: we restrict
ourselves to two fidelity-preserving interventions with no literature
risk: (i) removal of redundant content, and (ii) externalization of
content to permanent external documentation, replaced in-prefix by a
one-line pointer. No section is rewritten or compressed; only
unnecessary duplication is removed and stable content is moved to
dedicated external homes where it already belongs.

We apply this algorithm to \blackrim's \claudemd across three waves of
revision. The conservative approach removes 249 lines (35.2\,\%)
and 9\,419 characters (19.4\,\%), measurably reducing static-prefix
cost while maintaining instruction-following fidelity across the
post-trim agent runs. The second wave alone delivered a 21.5\,\% line
reduction (152 lines) and 16.9\,\% character reduction (8\,178 characters).

\subsection*{Contributions}

This work makes four contributions:

\begin{enumerate}
  \item \textbf{Cost model.} A formal cost-minimization model decomposing
    static-prefix loading under cache hit rate $\cachehit$ and per-section
    access frequency, yielding a decision rule for in-prefix residency
    (\cref{eq:decision}).
  \item \textbf{Conservative algorithm.} A trim procedure restricted to
    redundancy removal and externalization, deferring lossy compression
    until measurement bridges the instruction-fidelity gap (\cref{sec:algo}).
  \item \textbf{Measured deltas.} Application to \blackrim across three
    committed waves, confirming 35.2\,\% line reduction and 19.4\,\%
    character reduction without observed instruction-fidelity degradation
    (\cref{sec:eval}).
  \item \textbf{Reproducibility artifacts.} Full data pipeline from live
    repository state, enabling refresh of results as the codebase evolves
    (\cref{sec:repro}).
\end{enumerate}

\subsection*{Paper roadmap}

\cref{sec:related} surveys prompt compression, hierarchical context design,
caching architectures, and multi-agent coordination systems, situating our
approach relative to adjacent work. \cref{sec:problem} formalizes the
cost-minimization problem and derives the core decision rule. \cref{sec:system}
describes \blackrim's architecture and how static prefixes are instantiated in
practice. \cref{sec:algo} presents the conservative trim algorithm and
justifies its restrictions. \cref{sec:eval} reports empirical measurements
and reproducibility setup. \cref{sec:discussion} surveys ten further
optimisation directions and discusses which are compatible with static-prefix
architecture. \cref{sec:future} identifies open questions and next steps.
