\section{Related Work}
\label{sec:related}

\subsection{Prompt compression and context reduction}

Aggressive prompt compression is well-established in retrieval contexts.
LLMLingua and LongLLMLingua use token importance scoring to compress
retrieval passages and question-answer pairs, reducing context size while
preserving downstream task performance \cite{llmlingua,longllmlingua}.
RECOMP improves retrieval-augmented language models through selective
augmentation and compression of retrieved passages \cite{recomp}.
AutoCompressors adapt language models themselves to compress contexts
\cite{autocompressors}, and ICAE applies in-context autoencoders for
similar purposes \cite{icae}. All these methods report maintained or improved
performance on standard QA and retrieval benchmarks. However, a critical
measurement gap exists: none benchmark compression specifically on instruction
documents. Instruction-following is qualitatively different from passage
retrieval (where only the answer matters), involving procedural reasoning,
rule application, and constraint satisfaction across the entire document.
The fidelity preservation guarantees from retrieval-context compression do not
automatically transfer to instruction compression, creating the gap our
conservative methodology navigates.

\subsection{Hierarchical and progressive context design}

MemGPT's three-tier memory model (RAM, recall, and archival storage) pioneered
a frequency-weighted approach to context residency, motivated by human working
memory and the desire to support long-horizon tasks \cite{memgpt}. The key
insight—that not all information needs resident in-context at once—directly
parallels our frequency-weighted disclosure model. However, MemGPT operates at
the episodic-memory level (conversation history), whereas our model addresses
the procedural-instruction level (static prefixes). Liu et al.'s ``Lost in the
Middle'' study empirically established that language models attend
preferentially to early and late positions in long contexts \cite{lostinmiddle}.
Importantly, the U-shaped degradation curve is pronounced for retrieval tasks
(15--25\,\% quality loss in the middle) but modest for instruction-tuned models
(3--8\,\% only). This finding validates the structural positioning of static
instruction prefixes at context start as cost-optimal and motivates retention of
early-position stable rules.

\subsection{Prompt caching and KV-cache architecture}

Anthropic's prompt caching feature reduces token costs for repeated prefixes by
caching KV states within a 5-minute window at approximately 10\,\% of the miss
cost \cite{anthropic-cache}. The architecture is the operational baseline we
assume throughout this work. ``Don't Break the Cache,'' an independent
evaluation, documents failure modes and cost gotchas in long-horizon agentic
tasks \cite{dont-break-cache}. At the infrastructure layer, PagedAttention
enables efficient KV-cache memory management in open-source serving systems
\cite{pagedattention}, and distributed systems such as Mooncake and Preble
extend caching to multi-server deployment \cite{mooncake,preble}. Our cost
model in \cref{sec:problem} subsumes these caching properties as a parameter
$\cachehit$; the progressive-disclosure algorithm applies regardless of whether
the cache is Anthropic's or open-source.

\subsection{Knowledge graphs and structured augmentation}

GraphRAG and KAG propose knowledge-graph structures to organize and retrieve
relational information, designed for unstructured corpora and business
intelligence contexts \cite{graphrag,kag}. Both emphasize discovering implicit
relationships and enabling discovery queries over large document collections.
However, instruction documents are fundamentally different: they are procedural,
not relational; the instruction-following task is to apply rules correctly, not
to discover hidden relationships. Knowledge graphs introduce serialization
overhead (nodes and edges must be encoded into text) that dominates gains in
procedural domains. We therefore argue these approaches, while valuable for
knowledge work, are misaligned with the static-instruction problem and do not
apply to the conservative trim.

\subsection{Multi-agent coding systems}

AutoGen, MetaGPT, and ChatDev pioneered agent coordination architectures that
dispatch specialized agents across roles, leveraging shared memory and
communication patterns \cite{autogen,metagpt}. These systems motivated the
multi-agent static-prefix problem: when each agent spawn loads the same
instruction document, the total context cost becomes a critical bottleneck at
scale. Our contribution sits beneath the coordination layer, optimizing the
per-agent prefix residency. Notably, the empirical instantiation in this paper
(the \blackrim multi-agent framework) makes dogfooding measurement possible:
the static prefix itself documents the instruction dispatch pattern, enabling
closed-loop frequency analysis. This measurement opportunity is unique to coding
systems where agents follow programmatic, statically-analyzable dispatch rules.
