\section{Problem Formulation}
\label{sec:problem}

\subsection{Setting}

Let $\prefix = (\section_{i})_{i=1}^{n}$ be a static instruction document
partitioned into $n$ sections. For each section we define:

\begin{itemize}
  \item $\size{\section_{i}}$ — size in tokens.
  \item $\frequency_i \in [0, 1]$ — fraction of agent spawns that
    semantically need this section's content.
  \item $\readsize_i$ — token cost of fetching $\section_{i}$ via an
    on-demand Read call (file content + tool-invocation overhead).
\end{itemize}

Let $\cachehit \in [0, 1]$ be the prompt-cache hit rate over a window
(empirically $0.8$--$0.9$ for Blackrim under 5-min TTL with $\sim$5
parallel-agent waves), and $S$ the number of spawns over the window.

\subsection{Cost model}

Without progressive disclosure (everything in-prefix), per-window cost
is

\begin{equation}
  \cost_{\mathrm{flat}} = \size{\prefix} \cdot
    \big( (1 - \cachehit)\,S + \cachehit\,S \cdot \alpha \big)
  \label{eq:cost-flat}
\end{equation}

where $\alpha \approx 0.10$ is the cache-read multiplier (per
\cite{anthropic-cache}). Cache misses pay full price; hits pay
$\sim 10\%$.

With frequency-weighted progressive disclosure, each section is either
\emph{resident} ($\section_{i} \in \prefix$) or \emph{externalised}
(\(\section_{i}\) lives in an external doc accessed via Read on demand).
Per-window cost becomes

\begin{equation}
  \cost_{\mathrm{trimmed}} =
    \!\!\sum_{i \in \mathrm{resident}}\!\! \size{\section_{i}} \cdot
      \big( (1 - \cachehit)\,S + \cachehit\,S\,\alpha \big)
    \;+\;
    \!\!\sum_{j \in \mathrm{ext}}\!\! \frequency_j \,S \cdot \readsize_j .
  \label{eq:cost-trimmed}
\end{equation}

\subsection{Decision rule}

A section $\section_{i}$ is more cost-efficient externalised iff

\begin{equation}
  \frequency_i \,\readsize_i < \size{\section_{i}}
    \cdot \big( (1 - \cachehit) + \cachehit \alpha \big) .
  \label{eq:decision}
\end{equation}

Substituting Anthropic's published $\alpha = 0.10$ and Blackrim's
empirical $\cachehit \approx 0.85$ gives a break-even when
$\frequency_i \cdot \readsize_i < 0.235\,\size{\section_{i}}$.

\subsection{Fidelity constraint}

Equation~\eqref{eq:decision} ignores quality. We add a
\emph{fidelity-preservation constraint}: a section may be externalised
only if its absence from the resident prefix does not measurably
degrade instruction-following quality on a held-out eval. Formally:
let $Q(\prefix')$ denote eval-suite quality given prefix $\prefix'$.
Then the trim is valid iff $Q(\prefix_{\mathrm{trimmed}}) \geq
Q(\prefix_{\mathrm{flat}}) - \epsilon$ for tolerance $\epsilon$.

\tothink{Empirical $\epsilon$ choice + held-out eval suite specification
deferred to \cref{sec:eval}; formal constraint handling in \cref{sec:algo}.}
