\section{System: Blackrim Multi-Agent Architecture}
\label{sec:system}

\subsection{Multi-agent architecture}

\blackrim is a multi-agent framework for software engineering tasks organized
as a crew of domain-specialized agents. The orchestrator (``Gestalt'') is a
single opus-class model that routes work to the appropriate citizen based on
domain. Eight citizens—each an opus-class model and domain lead (Engineering,
Frontend, Infrastructure, Security, Product, Data, ML, Knowledge)—own vertical
slices and dispatch workers for parallel or specialized work. Seven workers
(sonnet/haiku, shared pool) execute functional tasks: architecture review,
code building, testing, code review, research, writing, and scoring.

Coordination across agents happens via \texttt{bd} (beads), a Dolt-backed task
graph that assigns issue IDs, tracks dependencies, and survives context-window
resets. The dispatch pattern is atomic: scan ready tasks, claim one or more
tasks, spawn parallel agents in isolated worktrees, await completion, merge
results, and close tasks. Because multiple agents may spawn in parallel within a
single 5-minute prompt-cache window, the cost model in \cref{sec:problem}
reflects the empirical hit-rate structure observed during multi-agent waves.

\subsection{Static-prefix instruction document layers}

\blackrim's instruction architecture is three-tiered. A global \texttt{CLAUDE.md}
at \texttt{\~{}/.claude/} is thin and rarely changes; it loads automatically
into every Claude Code session across all projects. The project-level
\texttt{CLAUDE.md} (the focus of this paper) is thick and project-specific,
defining domain roles, dispatch discipline, skill definitions, and agent
coordination patterns. Per-agent overlays—\texttt{agents/citizens/*.md} and
\texttt{agents/workers/*.md}—are prepended to the base prefix at dispatch,
allowing agents to declare their specific capabilities and constraints.

The project \texttt{CLAUDE.md} is the single source of truth for instruction
codification. Its content directly affects agent behavior (which skills are
available, how dispatch routes work, what coordination patterns are expected),
making it a critical operational document. Consequently, static-prefix loading
cost is dominated by the project-level document, measured at 708 lines and
48\,487 characters at the baseline commit.

\subsection{Telemetry and measurement}

Every agent spawn in \blackrim logs to \texttt{.beads/telemetry/invocations.jsonl},
recording the spawn timestamp, agent role, task ID, input token count, output
token count, and crucially, Anthropic's cache-state fields:
\texttt{cache\_creation\_input\_tokens} (first spawn of a window) and
\texttt{cache\_read\_input\_tokens} (subsequent spawns). These fields enable
direct measurement of prompt-cache behavior under real workloads.

The evaluation in \cref{sec:eval} derives its measurements from git history
(CLAUDE.md size trajectory at commit-level granularity) paired with post-trim
agent execution traces (confirming fidelity through successful task completion).
The reproducibility setup (\cref{sec:repro}) documents the data pipeline that
regenerates measurements from live Blackrim repository state, enabling future
verification and refresh as the framework evolves.
