\section{Algorithm: Conservative Frequency-Weighted Trim}
\label{sec:algo}

\subsection{Inputs}

\begin{itemize}
  \item Current $\prefix$ partitioned into sections $(\sect)_{i=1}^{n}$.
  \item Per-section access-frequency estimates $\hat{\frequency}_i$ (from
    domain knowledge, telemetry, and the literature-grounded "what does
    every spawn need" heuristic).
  \item Existing or to-be-created external homes for externalised content.
  \item A held-out instruction-following eval suite for the fidelity check.
\end{itemize}

\subsection{The conservative trim algorithm}

\begin{algorithm}[H]
\SetAlgoLined
\KwIn{$\prefix$, $\{\hat{\frequency}_i\}$, $\cachehit$, $\epsilon$}
\KwOut{$\prefix_{\mathrm{trimmed}}$}
\For{each section $\sect$ in $\prefix$}{
  Classify $\sect$ as one of: \emph{must-resident},
    \emph{frequency-deciding}, or \emph{redundant}.\;
  \uIf{\emph{must-resident}}{Keep in $\prefix_{\mathrm{trimmed}}$.\;}
  \uElseIf{\emph{redundant}}{Drop or replace with one-line pointer.\;}
  \Else(\emph{frequency-deciding}){
    \If{Eq.~\eqref{eq:decision} prefers external}{
      Verify a permanent external home exists; if not, file a
        prerequisite to create one before this step.\;
      Replace section with one-line pointer to external doc.\;
    }
    \Else{Keep in $\prefix_{\mathrm{trimmed}}$.\;}
  }
}
Run held-out eval suite on $\prefix_{\mathrm{trimmed}}$;
  reject trim if $Q(\prefix_{\mathrm{trimmed}}) < Q(\prefix) - \epsilon$.\;
\Return{$\prefix_{\mathrm{trimmed}}$}\;
\caption{Conservative frequency-weighted trim}
\label{alg:trim}
\end{algorithm}

\subsection{Why \emph{conservative}}

The literature gap noted in \cref{sec:related} (no compression
benchmarks on instruction documents) means that learned compression
\emph{could} preserve fidelity but we have no measurement to back
it. Our algorithm therefore restricts itself to two interventions
that are provably fidelity-neutral:

\begin{enumerate}
  \item \textbf{Redundancy removal} — content that is duplicated
    within the prefix or duplicated against an existing canonical
    source elsewhere is dropped.
  \item \textbf{Externalisation with pointer} — content that exists
    permanently in an external doc is replaced by a one-line pointer.
    The LLM's tool-use is the retrieval mechanism.
\end{enumerate}

Both leave instruction-bearing content unchanged in its canonical
location. No section is rewritten to be terser; redundant statements
are simply removed.

\subsection{Worked examples from the Blackrim trim}
\label{sec:worked-examples}

\Cref{tab:section-residency} (Appendix) records per-section line counts at
baseline, Wave~1, and Wave~2 for the Blackrim \texttt{CLAUDE.md}.
Three rows illustrate the algorithm's three code paths:

\paragraph{Worktree Isolation — frequency-deciding, externalised.}
The \textit{Worktree isolation} sub-section of \textsection\,Background
Agents carried 76~lines at baseline (\emph{frequency-deciding}
classification).  Its content was already canonical in
\texttt{mkdocs/operations/worktree-guard.md}; the decision rule
Eq.\,\eqref{eq:decision} holds because the section is accessed only by
agents that are actively debugging an isolation failure, not by every
spawn.  After Wave~2 the sub-section collapsed to 14~lines — a single
paragraph plus a \texttt{Full reference:} pointer — an 82\% reduction
with no instruction content removed from its canonical home.

\paragraph{Customization — redundant, dropped.}
The \textit{Customization} section (28~lines at baseline) duplicated
content already maintained in \texttt{mkdocs/operations/customization.md}.
Because the section carried no must-resident directives not covered by
the external doc, it was classified \emph{redundant} and trimmed to a
compact 14-line summary by Wave~1, then held stable at Wave~2
(see \Cref{tab:section-residency}, row ``Customization'').

\paragraph{Delegation discipline — must-resident, kept.}
The \textit{Delegation discipline} section (44~lines) maps trigger
conditions to agent types.  Every Gestalt spawn consults it to decide
whether to self-do or delegate; $\hat{f} \approx 1$.  Applying
Eq.\,\eqref{eq:decision}: even at $\cachehit = 0.85$ and $\alpha = 0.10$
the break-even requires $\readsize < 0.235 \times 44 \approx 10$~lines,
far below the actual on-demand fetch cost.  The algorithm correctly keeps
it resident across all waves (0~lines removed,
\Cref{tab:section-residency}, row ``Delegation discipline'').

\Cref{fig:section-classification} summarises all nine top-level sections
simultaneously, showing how the algorithm partitions baseline line counts
across the three verdict classes.
\input{figures/fig2-section-classification}
