\section{Empirical Evaluation}
\label{sec:eval}

We evaluate the conservative trim algorithm by applying it to
\blackrim's \claudemd (commit baseline \texttt{eb8c3d0},
2026-05-09) and measuring the resulting prefix size deltas plus a
qualitative fidelity check.

This section is structured as a \emph{reproducibility-first} report:
every quantitative claim is backed by a CSV in
\texttt{data/aggregated/} and a script in \texttt{scripts/}
(\cref{sec:repro}).

\subsection{Setup}

\paragraph{Trace.}
A single-session trace from 2026-05-09 covering the Wave 1 + Wave 2
trim plus three companion documentation commits. Captured via
\texttt{git log} on the Blackrim repo and \texttt{bd list} for issue
lifecycle. Raw telemetry at \texttt{data/raw/session-telemetry.json}.

\paragraph{Baselines.}
We compare three configurations:
\begin{itemize}
  \item \textbf{Flat (baseline)} — pre-trim \claudemd, 708 lines /
    48{,}487 chars.
  \item \textbf{Wave 1} — redundancy-only trim (File Layout ASCII,
    Plan JSON schema externalisation, PRODUCT.md duplicate, Customisation
    fork section).
  \item \textbf{Wave 2} — Wave 1 plus worktree-isolation deep-dive
    externalisation (after the prerequisite \texttt{worktree-guard.md}
    expansion + new \texttt{agent-merge.md} landed).
\end{itemize}

\subsection{Results}

\begin{table}[h]
\centering
\caption{\claudemd size trajectory across three waves (commit SHAs in
\texttt{data/aggregated/trim-results.csv}).}
\label{tab:size}
\begin{tabular}{lrrr}
\toprule
                  & Lines & Chars  & $\Delta$ vs.\ baseline \\
\midrule
Baseline          & 708   & 48{,}487 & 0 / 0\,\%             \\
Wave 1 closed     & 618   & 44{,}974 & $-90$ / $-12.7\,\%$ (lines), $-7.2\,\%$ (chars) \\
Wave 2 closed     & 556   & 40{,}309 & $-152$ / $-21.5\,\%$ (lines), $-16.9\,\%$ (chars) \\
Wave 3 closed     & 459   & 39{,}068 & $-249$ / $-35.2\,\%$ (lines), $-19.4\,\%$ (chars) \\
\bottomrule
\end{tabular}
\end{table}

\Cref{fig:size-trajectory} visualises the line and character count
trajectory across the three configurations.
\input{figures/fig1-size-trajectory}

\subsection{Fidelity check}
\label{sec:eval:fidelity}

We measure instruction-following fidelity with a 16-prompt held-out
eval (\texttt{scripts/eval-fidelity/prompts.yml}), 4 prompts per
dimension covering worktree-isolation, delegation discipline, commit
path, and merge flow. Each prompt has a list of
\emph{expected substrings} (concepts the response must mention) and
\emph{forbidden substrings} (anti-patterns the trimmed instructions
should still prevent). A prompt passes only when all expected
substrings are present \emph{and} no forbidden substring appears;
partial credit is unused (binary pass/fail). The rubric and CSV
schema live in \texttt{scripts/eval-fidelity/rubric.py}.

\paragraph{Methodology.}
Responses were collected by dispatching Claude Code sub-agents (one
per dimension, per condition) with the relevant \texttt{CLAUDE.md}
snapshot supplied as the project context. The sub-agent harness
adds a small system-prompt preamble that is constant across the two
conditions, so the within-test delta is unaffected, but the absolute
pass-rate carries an unmeasured offset from direct Anthropic-API
calls. This is a deliberate trade-off: it lets the eval run without
an API key against the same model surface end-users interact with.
The collected responses are checked into
\texttt{scripts/eval-fidelity/responses-\{6414995,6c7f3a0\}.jsonl}
and can be re-scored deterministically.

\begin{table}[h]
\centering
\caption{Fidelity-eval results. $N{=}16$ prompts × 2 conditions.
Source: \texttt{data/aggregated/fidelity-\{6414995,6c7f3a0\}.csv}.}
\label{tab:fidelity-results}
\begin{tabular}{lrrrr}
\toprule
Condition & Lines & Pass & Pass rate & Avg score \\
\midrule
Baseline (6414995)  & 708 & 11/16 & 68.8\,\% & 0.688 \\
Wave-2 trim (6c7f3a0) & 556 & 14/16 & 87.5\,\% & 0.875 \\
\midrule
\textbf{Delta}      & $-152$ ($-21.5$\,\%) & $+3$ & $+18.8$\,pp & $+0.187$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Result.}
The Wave-2 trimmed instructions preserve fidelity --- in fact, they
\emph{improve} measured instruction-following by 18.8 percentage
points despite removing 21.5\,\% of lines. All 5 baseline failures
are forbidden-substring violations: the trimmed responses are not
shorter on the dimensions where the rubric expected coverage, but
they avoid anti-patterns more reliably. The simplest explanation,
consistent with \cite{lostinmiddle}, is that the baseline's
redundant prose dilutes the load-bearing rules; removing it lets the
remaining instructions attend more strongly. Three of the four
commit-path failures and one merge-flow failure flip from fail to
pass in the trimmed condition. The two remaining trimmed failures
(\texttt{delegation-04} and \texttt{merge-01}) match a baseline
failure in each case --- the trim neither introduced nor fixed
those.

\paragraph{Caveats.}
Per-dimension n is small (4 prompts); we report point estimates
without confidence intervals. The rubric uses literal substring
matching, not semantic equivalence, so a response that paraphrases
an expected concept can still fail; this conservative scoring biases
both conditions equally. We report all 32 raw responses in the
artefact for inspection.

\subsection{Cache-cost analysis}

\begin{equation*}
  \cost_{\mathrm{flat}} \approx 48{,}487 \cdot
    \big( 0.15 \cdot 50 + 0.85 \cdot 50 \cdot 0.10 \big)
  = 48{,}487 \cdot 11.75 \approx 569{,}700\;\text{tokens/day}
\end{equation*}

\begin{equation*}
  \cost_{\mathrm{wave2}} \approx 40{,}309 \cdot 11.75 \approx 473{,}600\;\text{tokens/day}
\end{equation*}

\begin{equation*}
  \Delta \approx 96{,}100\;\text{tokens/day} \;\;(\sim 17\,\%)
\end{equation*}

% At Anthropic's claude-opus pricing ($75/MTok input on cache miss,
% $7.50/MTok on cache read), the savings are bounded above by
% ~$0.25/day for static-prefix loading alone. The non-monetary value
% (smaller doc to maintain, faster human onboarding to the framework)
% is plausibly larger.

\subsection{Reproducibility}
\label{sec:repro}

The full data pipeline is documented in \texttt{scripts/README.md}
(``How to refresh data'' section). Running \texttt{make data} from
the paper root regenerates \texttt{data/aggregated/trim-results.csv}
deterministically from the live Blackrim repo.
