\section{Discussion}
\label{sec:discussion}

\subsection{Why conservative beat learned compression}

The most striking finding from this work is that fidelity-neutral interventions
(redundancy removal and externalisation) delivered the majority of the token
gains without any learned compression machinery. Our 21.5\,\% line reduction
and 17.1\,\% cache-amortised token saving came exclusively from removing
duplication and relocating infrequently-needed procedural content to external
documents.

This result aligns with a structural property of instruction documents: they
accumulate redundancy by design. A framework like \blackrim's \claudemd serves
both as a reference manual and a living operational guide. Authors naturally
explain concepts multiple times (in Customisation sections, in agent-specific
guidance, in glossaries) to ensure readers find answers at hand without
hunting. Once this duplication is visible via frequency analysis, removing it
causes no fidelity loss.

Literature on learned compression (LLMLingua, RECOMP, AutoCompressors) focuses
on retrieval-context reduction: summarising QA pairs and news articles where
the task is to extract a single relevant fact. Instruction documents are
procedurally authored. Critical sections like worktree-isolation rules and
commit paths are written twice (once in deep-dive prose, once in a checklist)
not by accident but because different readers consume them differently: some
scan the checklist at-a-glance, others want the full narrative. A learned
compressor trained on retrieval corpora would see both as noise and risk dropping
either, degrading fidelity. Our algorithm makes the conservative choice: it
respects the author's redundancy as a signal of what matters.

\subsection{Generalisability}

The 21.5\,\% line reduction and 17.1\,\% token saving measured on
\blackrim's \claudemd are encouraging but case-study-specific. Will the
algorithm transfer to other multi-agent systems?

Transfer depends on three signals:

\begin{enumerate}
  \item \textbf{Prefix size.} Systems with larger static prefixes (100 KB+)
    see more absolute token savings. Systems with minimal prefixes (<10 KB)
    may not justify the audit effort. \blackrim's 49 KB baseline puts it
    in the high-impact regime.
  \item \textbf{Duplication rate.} The trim's gains are proportional to
    redundancy present in the original document. A carefully-curated
    single-file instruction document with no duplication will see
    minimal extraction. \blackrim's ~5 year history and multi-citizen
    governance accumulated duplication intentionally; younger systems
    may see half the gains.
  \item \textbf{External-home availability.} Externalisation only works
    if stable external documents exist to point to. \blackrim benefits
    from a mature mkdocs site and a design-decision archive (ADR path).
    Systems with monolithic instruction sets and no reference web have
    fewer extraction targets.
\end{enumerate}

Practitioners evaluating the algorithm should audit their own instruction
documents against these three signals. A simple pre-assessment pass (scan
for duplicate paragraphs, count unique external references, estimate
cache-TTL-amortised token impact) takes 1--2 hours and predicts ROI
before committing to a full trim.

\subsection{Limitations}

Four limitations bound the strength of this work:

\begin{enumerate}
  \item \textbf{Single-system case study.} This is a measurement of one
    multi-agent framework applied to its own instruction document. No
    held-out systems are evaluated. The generalisability argument above
    is qualitative, not validated empirically.
  \item \textbf{Eval-suite dimension coverage.} The fidelity check in
    \cref{sec:eval} is qualitative: agents completed successfully post-trim.
    A held-out, adversarial instruction-following eval suite (probing
    worktree-isolation rules, delegation thresholds, commit discipline,
    merge-flow correctness) would be substantially more rigorous. The work
    notes this as future work but does not discharge it.
  \item \textbf{Dynamic cost unmeasured.} This paper measures static
    prefix size and amortised token cost under the Anthropic prompt-cache
    model (80\,\% hit rate, 5-minute TTL). It does not measure the latency
    cost of dynamic disclosure: when an agent calls \texttt{read\_doc}
    or similar to fetch an externalised section on-demand, what is the
    extra RPC round-trip cost? The token savings may be offset by
    interaction latency.
  \item \textbf{Cache-hit rate assumed constant.} The analysis assumes
    $\rho = 0.85$ (cache hit rate) is invariant across prefix sizes.
    In practice, more frequent cache invalidation or reduced prefix
    stability might degrade $\rho$. We do not model this sensitivity.
\end{enumerate}

\subsection{Threats to validity}

The most significant threat is the measurement setup itself. The
\blackrim repository is the primary author's daily-driver system. The
frequency estimates, section classifications, and external-home decisions
were made with deep domain knowledge of how agents actually use the
framework. When a new maintainer applies the algorithm cold---without
that embedded understanding---classification decisions will differ. Some
sections externalised here might be kept in-prefix by a reader unfamiliar
with Blackrim's workflow; other sections might be externalised too
aggressively.

This is not a flaw in the algorithm (it is explicitly conservative and
includes an eval gate); it is a ceiling on the claim. The 21.5\,\% figure
is reproducible only if the frequency estimates and classification judgments
are accepted as-is. A second independent practitioner applying the
same algorithm to the same document would likely arrive at 18--25\,\% savings
depending on their own usage patterns and risk tolerance.
