\begin{abstract}
\noindent
In a static-prefix multi-agent LLM coding system with a well-tuned
subagent dispatch tier---16 static rules routing each subagent role
to haiku, sonnet, or opus---dispatched subagents account for only
\textbf{0.6\,\%} of total session cost. The orchestrator's main
thread accounts for the remaining \textbf{99.4\,\%}. Once dispatch
is tuned, further dispatch optimisation yields negligible savings;
the orchestrator itself is the binding cost lever. We present two
complementary primitives that target this lever.

The first is a \emph{semantic-similarity per-turn router}: each
main-thread turn is embedded and matched against a hand-curated
exemplar set (\(\sim\)150 labeled turns, 50 per tier); the top
nearest neighbour's tier label is returned, with conservative
escalation to opus when no exemplar clears a similarity threshold
\(\tau\). Evaluated via leave-one-out cross-validation on a 49-turn
held-out test set drawn from real \blackrim session transcripts, the
router achieves macro-F1 0.526 (+6.0 points over the
length-heuristic baseline), haiku-F1 0.786, and estimated per-batch
cost savings 30\,\% above the baseline.

The second is an \emph{agentic plan-level semantic cache}, following
\citeauthor{plancache}~\cite{plancache}: plans for structurally
equivalent bd dispatches are keyed on a stable semantic signature
(role, task shape, normalised file scope, deliverable kind) and
their outputs cached with LRU eviction and a 30-day TTL. The cache
stacks atop the existing prefix cache rather than replacing it.
We characterise the calibration baseline: a 10.44\(\times\)
read-to-creation ratio, 90.1\,\% hit rate, and 78.6\,\% estimated
savings on real-API spawns from Anthropic's \texttt{cache\_control}
layer. Plan caching targets the remaining 21.4\,\% by capturing
reuse across the 5-minute prefix-cache TTL window.

Per-turn routing and plan caching are independent in the cost
equation; their savings compound multiplicatively. We derive a
closed-form composed-cost bound and document an integration
blocker---the current Claude Code interface exposes session-level
model selection only, not per-turn routing---with three evaluated
workaround paths. All evaluation data, scripts, and labels are
released to support reproducibility.
\end{abstract}
