\section{Related Work}
\label{sec:related}

This paper sits at the intersection of two bodies of work: LLM model
routing (choosing which model to invoke given a query) and caching for
LLM systems (avoiding redundant computation). We survey both strands,
noting where each applies and where it falls short for the per-turn
multi-agent orchestration setting we address.

\subsection{Model routing and cascading}
\label{sec:related-routing}

\paragraph{FrugalGPT.}
\citeauthor{frugalgpt}~\cite{frugalgpt} introduce the
\emph{cascade} framing: invoke a cheap model first and escalate
to a more capable (and expensive) model only when the cheap
model's output fails a quality check. On five standard NLP
benchmarks they demonstrate 98\,\% cost savings at 95\,\%+
fidelity relative to always using GPT-4. The key idea---that a
quality signal on the cheap model's output can gate escalation---
applies in principle to per-turn orchestration. In practice,
however, the gate requires an observable quality signal
\emph{within the same turn}, before the downstream subagent
dispatches have run; in our setting the turn's quality is
not fully observable until downstream execution completes, making
synchronous cascade impractical. We adopt conservative escalation
(route to opus when uncertain) as the operationally safe
analogue.

\paragraph{RouteLLM.}
\citeauthor{routellm}~\cite{routellm} train a lightweight
classifier on human preference data (\textsc{ChatBot Arena}
labels) to distinguish which model tier a given prompt requires.
Their best system achieves 85\,\% cost reduction on MT-Bench at
95\,\% of GPT-4 quality. RouteLLM's core insight---that a small
trained router can recover most of a strong model's quality at
a fraction of the cost---is the inspiration for our exemplar-based
router, but preference-data training is not available for
main-thread orchestration turns: there is no pairwise human
preference signal for ``Gestalt's response to a bd-dispatch prompt
on haiku vs.\@ opus.'' We therefore use exemplar curation rather
than supervised training for the v1 router, deferring the trained
classifier to future work.

\paragraph{Hybrid LLM and Tryage.}
\citeauthor{hybridllm}~\cite{hybridllm} route queries between a
small local model and a large remote model by estimating
\emph{output quality} from the query alone, without observing the
small model's output first. \citeauthor{tryage}~\cite{tryage}
extend this to real-time routing across an arbitrary catalogue of
models by learning capability-aware representations for each
provider. Both systems target single-turn question-answering
workloads; neither accounts for the rolling conversation context
or the structured-tool-use success signal available in a
multi-agent session. Our router incorporates a rolling \(K\)-turn
context window to capture conversational momentum, which is
critical for disambiguating orchestration turns that look similar
in isolation (e.g., a short planning question and a short
confirmation message).

\paragraph{Confidence-based escalation.}
\citeauthor{semanticentropy}~\cite{semanticentropy} route queries
by measuring semantic entropy in a cheap model's output: high
entropy implies the cheap model is uncertain, triggering
escalation. \citeauthor{aggarwal2024routingconfidence}~\cite{aggarwal2024routingconfidence}
use speculative token-level confidence for a similar purpose.
Confidence-based routing has the attractive property of
\emph{self-certifying}: the cheap model's own distribution signals
whether it is adequate. However, it requires a first pass on the
cheap model for every turn, incurring non-trivial latency overhead
(a full haiku or sonnet inference call before deciding to route to
opus). In a long-running orchestration session where the operator
is actively waiting, this latency is unacceptable on turns that
will escalate anyway. Our approach classifies \emph{before}
inference from the operator's prompt alone, at the cost of a
single embedder forward pass.

\paragraph{What is missing for the multi-agent orchestration setting.}
The routing literature addresses either (a) subagent dispatch---
where the agent's role provides a strong structural prior and a
clean (role~\(\times\)~signal) rule cascade suffices
\cite{moa1landscape}---or (b) single-turn chat---where the
conversation has no persistent session state and tool-use success
is not observable. The main-thread orchestration problem is
neither: turns are drawn from a long-running session with shared
context, there is no role label, and the quality signal is
partially observable through downstream tool-use outcomes. No
published routing system is grounded in this setting, and no
evaluation data set of labeled main-thread orchestration turns
exists in the literature. This paper provides the first such
labeled set (49 turns, three-tier labels) and the first published
evaluation on it.

\subsection{Caching for LLM systems}
\label{sec:related-caching}

\paragraph{Prefix caching.}
Modern LLM serving systems cache the key-value (KV) attention
states for repeated prompt prefixes, avoiding redundant
computation on tokens that have been processed before.
\citeauthor{pagedattention}~\cite{pagedattention} introduce
PagedAttention for efficient KV-cache memory management in vLLM;
\citeauthor{mooncake}~\cite{mooncake} extend this to a
disaggregated prefill-decode architecture that decouples KV
cache from the compute layer. On the API side, Anthropic's
\texttt{cache\_control} headers~\cite{anthropic-cache} expose
prompt caching directly to client applications: cache-write
tokens are billed at 1.25\(\times\) the base rate, and cache-read
tokens at 0.10\(\times\)---a 12.5\(\times\) read-to-write ratio
that yields substantial savings when a prompt prefix is reused
within the 5-minute TTL window. \citeauthor{dont-break-cache}~\cite{dont-break-cache}
characterise the fragility of this savings: even small prefix
perturbations across spawns---reordered tool lists, timestamp
updates, non-deterministic field ordering---can break the cache
and eliminate the discount. \blackrim already achieves a
10.44\(\times\) read-to-creation ratio and 90.1\,\% cache hit rate
under deliberate prefix-stabilisation discipline (\S\ref{sec:eval});
prefix caching is therefore the \emph{calibration baseline} atop
which we measure additional savings.

\paragraph{Agentic plan caching.}
\citeauthor{plancache}~\cite{plancache} introduce plan caching as
a complement to prefix caching for agentic systems. Their key
observation is that many agent dispatches reduce to
\emph{structurally equivalent} plans---the task shape, agent role,
and deliverable kind are the same even if the surface text differs.
By hashing a semantic plan signature and caching the orchestration
output keyed on that hash, the full plan-composition step can be
short-circuited on hit. They report a 50\,\% cost reduction and
27\,\% latency reduction on their agentic benchmark relative to
no-cache. This paper's plan cache (\S\ref{sec:plancache}) is a
direct implementation of their algorithm applied to \blackrim's
bd-dispatch pipeline; our contribution on this side is
\emph{empirical}, not algorithmic: we measure the hit rate and
incremental savings in a production multi-agent coding system,
and we characterise the interaction between plan-cache savings and
the prefix-cache baseline. We do not claim novelty for the
caching mechanism itself and are explicit about this lineage.

\paragraph{The gap this paper fills.}
Existing work on LLM caching addresses either the KV-cache
layer (prefix caching, PagedAttention) or the plan layer
(agentic plan caching) in isolation. No prior work measures the
\emph{composed cost model} combining prefix caching, plan caching,
and per-turn routing in a single production system, or
characterises how savings from each source interact. We show that
the three sources are independent in the cost equation and that a
closed-form product bound applies under mild independence
assumptions (\S\ref{sec:problem}). We also provide the first
comparison against the prefix-cache baseline in a real-deployment
context where cache-break discipline is an active operational
concern.
