\section{Algorithm: Agentic Plan Caching}
\label{sec:plancache}

\subsection{Approach}

Agentic plan caching~\cite{plancache} exploits the observation that
orchestration plans for structurally similar tasks share most of their
strategy-level content.  In \blackrim, the orchestrator emits a plan
before each subagent dispatch: the plan specifies agent role, task
shape (e.g., extract, refactor, scaffold, review), file scope, and
expected deliverable kind.  Two dispatches may differ in surface text
(different bd issue IDs, different file names) while being identical
at the structural level.  If the plan for the earlier dispatch was
successful, its output is a valid starting point for the later dispatch
with minimal adaptation.

The plan cache indexes these structural outputs by a
\emph{semantic signature} \(\plan\) (Eq.~\ref{eq:plansig}).  On a new
dispatch query, the cache is consulted before the plan-composition
LLM call fires; a hit short-circuits the \texttt{gt context budget}
/ \texttt{gt orchestrate plan} path entirely.  This is the sense in
which plan caching is orthogonal to prefix caching: prefix caching
discounts the static-prefix portion of LLM calls that do execute;
plan caching prevents certain LLM calls from executing at all.

\subsection{Plan signature design}

\begin{equation}
  \plan = h\bigl(
    \textit{role},\;
    \textit{task\_shape},\;
    \textit{normalised\_file\_scope},\;
    \textit{deliverable\_kind}
  \bigr)
  \label{eq:plansig}
\end{equation}

where \(h\) is a stable hash (SHA-256 of the canonicalised tuple).
The signature is intentionally coarser than the full plan text.
File names are normalised to their directory prefix
(\texttt{internal/dispatch/} rather than
\texttt{internal/dispatch/model.go}) to increase hit rate across
dispatches that touch different files in the same module.  Deliverable
kinds are one of a closed vocabulary: \texttt{path-list},
\texttt{code-edit}, \texttt{prose}, \texttt{analysis},
\texttt{test-suite}.

\paragraph{The abstraction-level trade-off.}
Coarser signatures (fewer fields, broader normalisation) increase the
hit rate but also increase the false-hit rate: the cache may return a
plan that is structurally similar but contextually wrong.  Finer
signatures (more fields, exact file names) decrease both rates.  For
the evaluation in §\ref{sec:eval}, we use the four-field signature
above.  The false-hit guard (similarity threshold \(\theta_{\text{hit}}\)
on the plan embedding, not just the signature hash) provides a second
filter: even when two dispatches hash to the same signature, we verify
that their plan embeddings are similar above threshold before returning
the cached output.

\subsection{Algorithm}

\begin{algorithm}[H]
\SetAlgoLined
\DontPrintSemicolon
\KwIn{query plan signature $\plan_q$, query embedding $\phi(\plan_q)$,
      cache $\cache$, similarity threshold $\theta_{\text{hit}}$,
      confidence threshold $\theta_{\text{conf}}$}
\KwOut{cached plan output $o$ or \textsc{miss}}
\BlankLine
\If{$\plan_q \in \cache$ and $\cache[\plan_q].\textit{confidence} \geq \theta_{\text{conf}}$}{
  \Return{$\cache[\plan_q].\textit{output}$}
  \tcp*[r]{exact signature hit}
}
$N \gets \bigl\{\plan' \in \cache :
  \cos\!\bigl(\phi(\plan_q),\, \phi(\plan')\bigr) \geq \theta_{\text{hit}}\bigr\}$
\tcp*[r]{semantic fallback}
\uIf{$N \neq \emptyset$}{
  $\plan^{*} \gets \arg\max_{\plan' \in N}
    \cos\!\bigl(\phi(\plan_q), \phi(\plan')\bigr)$\;
  \Return{$\cache[\plan^{*}].\textit{output}$}\;
}
\Else{
  \Return{\textsc{miss}}\;
}
\caption{Plan-cache lookup with exact-match and semantic fallback}
\label{alg:plancache}
\end{algorithm}

On a miss, the orchestrator runs the standard plan-composition path
(\texttt{gt context budget} / \texttt{gt orchestrate plan}) and inserts
\((\plan_q, o)\) into \(\cache\) upon successful completion, i.e.,
the dispatch closed with a success verdict.  Cache fills on
success-only avoids poisoning the cache with failed or partial plans.

\subsection{Invalidation policy}

This paper's evaluation uses the simplest possible invalidation policy:
\emph{no invalidation} within a replay evaluation run.  This is
appropriate for a controlled replay study where the codebase does not
change between dispatches.

For production deployment, three invalidation policies are natural:

\begin{enumerate}[noitemsep]
  \item \textbf{Time-based TTL.} Entries expire after \(T_{\text{ttl}}\)
    days (e.g., 30 days).  This handles gradual codebase drift but
    does not respond to schema changes.
  \item \textbf{Code-version keying.} Each cache entry is keyed by a
    composite of \(\plan\) and the HEAD commit SHA of the repo at fill
    time.  A new commit invalidates entries whose file scope intersects
    the changed paths.  This is the most precise policy but requires a
    changed-file index.
  \item \textbf{Schema-change invalidation.} When the agent role
    vocabulary or deliverable-kind vocabulary changes (e.g., a new
    worker type is added), the entire cache is cleared.
\end{enumerate}

We note these as future-work directions (§\ref{sec:future}).  In the
evaluation (§\ref{sec:eval}), the no-invalidation assumption is
appropriate: the 30 historical bd dispatches are replayed in their
original sequence against a fixed codebase snapshot.

\subsection{Composition with prefix caching}

The interaction between plan caching and prefix caching is the key
source of superadditive savings.  The two mechanisms operate at
different layers and on different cost sources:

\begin{itemize}[noitemsep]
  \item \textbf{Prefix caching} (Anthropic's prompt cache,
    ~\cite{anthropic-cache}) discounts cache-read tokens at \(0.10\times\)
    the base rate within a 5-minute TTL window.  Savings accrue on the
    static-prefix portion (\claudemd, citizen definitions) of LLM
    calls that do fire.  The measured baseline across
    \blackrim's production telemetry: 10.44\(\times\) read-to-creation
    ratio, 90.1\,\% hit rate, 78.6\,\% estimated cost savings
    (§\ref{sec:eval}).
  \item \textbf{Plan caching} fires \emph{before} the LLM call:
    a plan-cache hit bypasses the \texttt{gt context budget} briefing
    step, so the briefing-composition LLM call is never made.  There
    is no cache-creation token cost and no cache-read token cost ---
    the entire call cost is eliminated.
\end{itemize}

Formally, for a plan-cache-hit dispatch the total token cost is
\(\cost_{\text{lookup}} \approx 0\) (hash lookup + embedding
comparison), versus the baseline
\(\cost_{\text{compose}}(\plan) = \alpha_{\tier} \cdot |\text{ctx}(\text{brief})| +
\beta_{\tier} \cdot |\text{out}(\text{plan})|\).  The prefix cache
cannot reduce a cost that is not incurred.  This is why the composed
savings of routing + plan caching exceed the sum of their individual
savings: plan caching saves entire calls (not just portions of calls),
and those calls are the ones on which prefix-cache savings would have
been largest (briefing compositions are long and share a common static
prefix).
